---
title: "What motivates a transfer: player performance, transfer fee or rivalry?"
author: "Vignesh Thiruvel Selvam"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(worldfootballR)
library(future)
library(furrr)
library(dplyr)
library(data.table)
library(stringr)
library(httr)
library(readr)
library(fuzzyjoin)
library(geosphere)
library(survival)
#library(ggplot2)
```

## Getting League data for each season

We get league table data for the top 5 leagues from 2017/18 season to 2024/2025 season.

```{r}
league_path <- "~/sem3/research_project/rival_analysis/data/league_table.csv"
if (!file.exists(league_path)) {
  leagues <- c("England", "Spain", "Germany", "Italy", "France")
  seasons <- 2017:2024
  league_season_combos <- expand.grid(league = leagues, season = seasons)
  
  # Parallel fetching
  results <- future_pmap_dfr(league_season_combos, function(league, season) {
    tryCatch({
      matchday <- case_when(
        league == "Germany" ~ 34,
        league == "France" & season == 2019 ~ 27,
        league == "France" & season >= 2023 ~ 34,
        TRUE ~ 38
      )
      message(paste("Fetching:", league, season))
      df <- tm_matchday_table(country_name=league, start_year=season, matchday=matchday)
      df$season <- season
      df
    }, error = function(e) {
      message(paste("Error fetching", league, season, ":", e$message))
      return(NULL)
    })
  })
  fwrite(df, league_path)
}
```

## Rivalry

We define rivalry based on three conditions:

-   In the previous season, teams that finished two places above or below a specific team are said to be its rivals.

-   Teams that have an ELO of 100 points more or less than the specific team are said to be its rivals.

-   Rivalry is considered only for teams that belong to the same league.

```{r}

# 1) read
dt_raw <- fread("~/sem3/research_project/rival_analysis/data/league_table.csv")

dt_raw[, team := str_squish(squad)]
dt_raw[, season := as.character(season)]
dt_raw[, league := as.character(league)]
dt_raw[, pos := as.integer(rk)]

if ("matchday" %in% names(dt_raw)) {
  dt_final <- dt_raw[!is.na(pos), .SD[which.max(as.numeric(matchday))], 
                     by = .(country, league, season, team)]
} else {
  dt_final <- dt_raw[!is.na(pos), .SD[1], by = .(country, league, season, team)]
}
# Keep only needed cols
dt_final <- dt_final[, .(country, league, season, team, pos)]
dt_final[, team_low := tolower(team)]
dt_final[, country_low := tolower(country)]
dt_final[, league_low := tolower(league)]

# --- 3) Build interval table (y) and point table (x) with distinct column names ---
# y: intervals owned by a team (pos_low .. pos_high)
dt_y <- dt_final[, .(country, league, season,
                     pos_low = pos - 2L,
                     pos_high = pos + 2L,
                     team_y = team,
                     pos_y = pos)]
# x: points to be matched into intervals
dt_x <- dt_final[, .(country, league, season,
                     pos_x = pos,
                     team_x = team,
                     pos_orig = pos)]
# Create start/end for pts (required by foverlaps)
dt_x[, pos_x_start := pos_x][, pos_x_end := pos_x]

# Set keys including country/league/season then interval start/end
setkey(dt_y, country, league, season, pos_low, pos_high)
setkey(dt_x, country, league, season, pos_x_start, pos_x_end)

# --- 4) Overlap join: match points to intervals within same country/league/season ---
ov <- foverlaps(x = dt_x, y = dt_y,
                by.x = c("country", "league", "season", "pos_x_start", "pos_x_end"),
                by.y = c("country", "league", "season", "pos_low", "pos_high"),
                nomatch = 0L)

# --- 5) Select relevant columns, remove self-matches, deduplicate unordered pairs ---
# ov contains columns from x and y; pick team_x (point) and team_y (interval owner)
rivals <- ov[, .(country = country, league = league, season = season,
                 team1 = team_x, pos1 = pos_x,
                 team2 = team_y, pos2 = pos_y)]

# remove identical teams and keep unordered unique pair per (country,league,season)
rivals <- rivals[team1 != team2]
rivals[, team_a := pmin(team1, team2)]
rivals[, team_b := pmax(team1, team2)]
rivals[, pos_diff := abs(as.integer(pos1) - as.integer(pos2))]
rivals <- unique(rivals[, .(country, league, season, team_a, team_b, pos_diff)])
setorder(rivals, country, league, season, team_a, team_b)

# --- 6) Add lowercase keys and set key for fast lookup ---
rivals[, team_a_low := tolower(team_a)]
rivals[, team_b_low := tolower(team_b)]
rivals[, country_low := tolower(country)]
rivals[, league_low := tolower(league)]
setkey(rivals, country_low, league_low, season, team_a_low, team_b_low)

# --- 7) Helper lookup --
# Usage: is_rival("England", "Premier League", "Arsenal", "Chelsea", "2020/2021")
is_rival <- function(country_in, league_in, t1, t2, season_in) {
  cl <- tolower(str_squish(country_in))
  ll <- tolower(str_squish(league_in))
  a <- tolower(str_squish(t1)); b <- tolower(str_squish(t2))
  a1 <- ifelse(a < b, a, b); b1 <- ifelse(a < b, b, a)
  res <- rivals[J(cl, ll, season_in, a1, b1), nomatch = 0L]
  nrow(res) > 0L
}


# Save output
fwrite(rivals, "~/sem3/research_project/rival_analysis/data/season_rivalries.csv")
```

```{r}
rivals
```

### ELO

```{r}
# ---------------- PARAMETERS ----------------
seasons <- 2018:2025
snapshot_month_day <- "-09-01"
max_name_dist <- 0.18

# ---------------- PATHS (uses uploaded files in ~/sem3/research_project/rival_analysis/data) ----------------
league_csv <- "~/sem3/research_project/rival_analysis/data/league_table.csv"
manual_map_path <- "~/sem3/research_project/rival_analysis/data/manual_club_map.csv"
out_dir <- "~/sem3/research_project/rival_analysis/data"
out_rivals_csv <- file.path(out_dir, "rivalry_lookup.csv")
out_elo_raw_dir <- file.path(out_dir, "elo_snapshots")
dir.create(out_elo_raw_dir, showWarnings = FALSE, recursive = TRUE)

# ---------------- 1) load & prepare league table (dt_final) ----------------
dt_raw <- fread(league_csv)
dt_raw[, team := str_squish(squad)]
dt_raw[, country := as.character(country)]
dt_raw[, league := as.character(league)]
dt_raw[, season := as.character(season)]
dt_raw[, pos := as.integer(rk)]

if ("matchday" %in% names(dt_raw)) {
  dt_final <- dt_raw[!is.na(pos), .SD[which.max(as.numeric(matchday))], by = .(country, league, season, team)]
} else {
  dt_final <- dt_raw[!is.na(pos), .SD[1], by = .(country, league, season, team)]
}
dt_final <- dt_final[, .(country, league, season, team, pos)]
dt_final[, team_clean := tolower(str_squish(team))]
# Precompute season_start string on dt_final for lookups (e.g., "2017" from "2017/18")
dt_final[, season_start_str := sub("([0-9]{4}).*", "\\1", season)]

# ---------------- 2) fetch ClubElo snapshots (cached) ----------------
fetch_elo_for_date <- function(date_str) {
  destfile <- file.path(out_elo_raw_dir, paste0("clubelo_", date_str, ".csv"))
  if (file.exists(destfile)) {
    message("Using cached Elo snapshot: ", destfile)
    df <- read_csv(destfile, show_col_types = FALSE)
    setDT(df); return(df)
  }
  url <- paste0("http://api.clubelo.com/", date_str)
  message("Fetching from API: ", url)
  r <- httr::GET(url)
  if (httr::status_code(r) != 200) stop("ClubElo request failed for date ", date_str, " (status ", httr::status_code(r), ")")
  txt <- httr::content(r, as = "text", encoding = "UTF-8")
  df <- read_csv(txt, show_col_types = FALSE)
  setDT(df)
  setnames(df, tolower(names(df)))
  club_col <- grep("club|team|name", names(df), value = TRUE)[1]
  elo_col  <- grep("elo|rating", names(df), value = TRUE)[1]
  if (is.na(club_col) || is.na(elo_col)) stop("Unable to find club/elo columns in ClubElo CSV for date: ", date_str)
  df <- df[, .(club = get(club_col), elo = as.numeric(get(elo_col)))]
  df <- df[!is.na(elo) & !is.na(club)]
  df[, club_clean := tolower(str_squish(gsub("\\(.*\\)", "", club)))]
  df <- df[order(-elo)]
  df[, snapshot_date := date_str]
  write_csv(df, destfile)
  message("Saved snapshot: ", destfile)
  return(df)
}

# Fetch and cache all snapshots (only does network calls if missing)
elo_all <- rbindlist(lapply(seasons, function(y) {
  d <- paste0(y, snapshot_month_day)
  message("Fetch snapshot: ", d)
  Sys.sleep(0.5) # be kind to API
  fetch_elo_for_date(d)
}), use.names = TRUE, fill = TRUE)
message("Total Elo rows fetched (or cached): ", nrow(elo_all))
elo_all[, season_start := as.character(as.integer(substr(snapshot_date, 1, 4)) - 1L)]

# ---------------- 3) APPLY MANUAL MAPPINGS (rowwise, robust) ----------------
manual_map_applied <- data.table()  # result holder
if (file.exists(manual_map_path)) {
  mm <- fread(manual_map_path)
  # normalize columns: support snapshot_date or season_start
  if (!("snapshot_date" %in% names(mm)) & ("season_start" %in% names(mm))) {
    mm[, snapshot_date := paste0(season_start, snapshot_month_day)]
  }
  if (!("club" %in% names(mm)) | !("mapped_team" %in% names(mm))) {
    stop("manual_club_map.csv must contain at minimum: club, mapped_team, and snapshot_date or season_start")
  }
  mm[, club_clean := tolower(str_squish(gsub("\\(.*\\)", "", club)))]
  mm[, mapped_team_clean := tolower(str_squish(mapped_team))]
  mm[, season_start := as.character(substring(snapshot_date, 1, 4))]
  # rowwise validate and build manual_map_applied
  valid_rows <- list()
  for (i in seq_len(nrow(mm))) {
    row <- mm[i]
    ss <- as.character(row$season_start)
    mt_clean <- row$mapped_team_clean
    # check dt_final: does mapped team exist in that season_start?
    cand <- dt_final[team_clean == mt_clean & season_start_str == ss]
    if (nrow(cand) == 0L) next  # invalid mapping: mapped_team not present in league for that season
    # find matching elo rows for the same snapshot + club_clean
    erows <- elo_all[season_start == ss & club_clean == row$club_clean]
    if (nrow(erows) == 0L) next  # no matching elo row to attach
    # build manual match rows
    for (er in seq_len(nrow(erows))) {
      for (cd in seq_len(nrow(cand))) {
        valid_rows[[length(valid_rows) + 1]] <- data.table(
          snapshot_date = erows[er, snapshot_date],
          club = erows[er, club],
          club_clean = erows[er, club_clean],
          elo = erows[er, elo],
          team = cand[cd, team],
          team_clean = cand[cd, team_clean],
          country = cand[cd, country],
          league = cand[cd, league],
          dist = 0  # mark manual mapping as exact
        )
      }
    }
  }
  if (length(valid_rows) > 0) {
    manual_map_applied <- rbindlist(valid_rows, use.names = TRUE, fill = TRUE)
    message("Applied ", nrow(manual_map_applied), " manual mapping rows from: ", manual_map_path)
  } else {
    message("No valid manual mappings found in: ", manual_map_path)
  }
} else {
  message("No manual_club_map.csv found at: ", manual_map_path, " (continuing without manual mappings)")
}

# ---------------- 4) FUZZY MATCH remaining Elo clubs -> dt_final teams ----------------
# Build candidate list excluding manual-mapped (by snapshot_date+club_clean)
elo_to_match <- copy(elo_all[, .(snapshot_date, club, club_clean, elo, season_start)])
if (nrow(manual_map_applied) > 0) {
  keys_mm <- unique(manual_map_applied[, .(snapshot_date, club_clean)])
  setkey(elo_to_match, snapshot_date, club_clean)
  setkey(keys_mm, snapshot_date, club_clean)

  # Proper anti-join (no need for fsetdiff)
  elo_to_match <- elo_to_match[!keys_mm]
}

# Build teams by season_start
teams_by_start <- unique(dt_final[, .(season_start_str, team, team_clean, country, league)])
setnames(teams_by_start, "season_start_str", "season_start")

matched_list <- list()
unmatched_list <- list()
for (yr in sort(unique(elo_to_match$season_start))) {
  eloy <- elo_to_match[season_start == yr, .(snapshot_date, club, club_clean, elo)]
  targ <- teams_by_start[season_start == yr, .(team, team_clean, country, league)]
  if (nrow(targ) == 0L) { unmatched_list[[yr]] <- eloy; next }
  mj <- stringdist_inner_join(eloy, targ, by = c("club_clean" = "team_clean"),
                              method = "jw", max_dist = max_name_dist, distance_col = "dist")
  setDT(mj); setorder(mj, snapshot_date, club, dist)
  if (nrow(mj) > 0) {
    mj <- mj[, .SD[1], by = .(snapshot_date, club)]  # best match per elo club
    matched_list[[yr]] <- mj
    matched_clubs <- unique(mj$club)
    unmatched_list[[yr]] <- eloy[!club %in% matched_clubs]
  } else {
    unmatched_list[[yr]] <- eloy
  }
}

matched_dt_fuzzy <- if (length(matched_list) > 0) rbindlist(matched_list, use.names = TRUE, fill = TRUE) else data.table()
unmatched_dt <- if (length(unmatched_list) > 0) rbindlist(unmatched_list, use.names = TRUE, fill = TRUE) else data.table()

# Combine manual-applied matches and fuzzy matches into matched_dt
if (nrow(manual_map_applied) > 0 && nrow(matched_dt_fuzzy) > 0) {
  matched_dt <- rbindlist(list(manual_map_applied, matched_dt_fuzzy[, .(snapshot_date, club, club_clean, elo, team, team_clean, country, league, dist)]),
                          use.names = TRUE, fill = TRUE)
} else if (nrow(manual_map_applied) > 0) {
  matched_dt <- copy(manual_map_applied)
} else {
  matched_dt <- copy(matched_dt_fuzzy)
}

# Save remaining unmatched for inspection
if (nrow(unmatched_dt) > 0) {
  fwrite(unique(unmatched_dt), file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
  message("Wrote remaining unmatched Elo clubs to: ", file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
} else {
  # produce empty file to indicate nothing remaining
  fwrite(data.table(snapshot_date=character(), club=character(), club_clean=character(), elo=numeric()),
         file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
  message("No remaining unmatched Elo clubs; created empty file: ", file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
}

# ---------------- 5) set Elo-season = snapshot_year - 1 (numeric) for Elo outputs ONLY ----------------
if (nrow(matched_dt) == 0) {
  message("Warning: matched_dt is empty (no Elo->league matches). Elo-only will be empty.")
}
matched_dt[, snapshot_year := as.integer(substr(snapshot_date, 1, 4))]
matched_dt[, elo_season := snapshot_year - 1L]  # e.g., 2018 -> 2017
```

### Rivalry Vector

```{r}

# ------Prepare points history-----
# Extract points table
pts_dt <- copy(dt_raw)

# normalize columns
pts_dt[, team_clean := tolower(str_squish(squad))]
pts_dt[, pts := as.numeric(pts)]

# keep only valid rows
pts_dt <- pts_dt[!is.na(pts),
                 .(country, league, season_start = season, team_clean, pts)]
pts_dt[, season_start := as.integer(season_start)]

# -------Build lagged points-------
# current season points
pts_t <- pts_dt[,.(season_start, team_clean, pts)]
setnames(pts_t, "pts", "pts_t")

# t-1
pts_t1 <- pts_dt[,.(season_start, team_clean, pts)]
pts_t1[, season_start := season_start + 1]
setnames(pts_t1, "pts", "pts_t1")

# t-2
pts_t2 <- pts_dt[,.(season_start, team_clean, pts)]
pts_t2[, season_start := season_start + 2]
setnames(pts_t2, "pts", "pts_t2")


# -----assemble base rivalry------

rival_base <- unique(
  pts_dt[, .(country, league, season_start, team_clean)]
)

# merge points
rival_base <- merge(rival_base, pts_t,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- merge(rival_base, pts_t1,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- merge(rival_base, pts_t2,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)


# ------attach elo-------
elo_dt <- matched_dt[
  , .(season_start = elo_season,
      team_clean, elo)
]

rival_base <- merge(rival_base, elo_dt,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- unique(rival_base, by = c("team_clean", "season_start"))


# ------- Impute missing POINTS using league-season quartile ------------

# compute league-season quartiles
pts_quartiles <- rival_base[
  , .(
      q_pts_t1 = quantile(pts_t1, 0.25, na.rm = TRUE),
      q_pts_t2 = quantile(pts_t2, 0.25, na.rm = TRUE)
    ),
  by = .(country, season_start)
]

rival_base <- merge(rival_base, pts_quartiles,
                    by = c("country", "season_start"),
                    all.x = TRUE)

# impute
rival_base[is.na(pts_t1), pts_t1 := q_pts_t1]
rival_base[is.na(pts_t2), pts_t2 := q_pts_t2]

# ------ Impute missing ELO using season quartile -------

elo_quartiles <- rival_base[
  , .(q_elo = quantile(elo, 0.25, na.rm = TRUE)),
  by = .(season_start)
]

rival_base <- merge(rival_base, elo_quartiles,
                    by = "season_start",
                    all.x = TRUE)

rival_base[is.na(elo), elo := q_elo]


# flag seasons where lagged points are structurally missing
lag_availability <- rival_base[
  , .(
      has_pts_t1 = any(!is.na(pts_t1)),
      has_pts_t2 = any(!is.na(pts_t2))
    ),
  by = .(country, season_start)
]

rival_base <- merge(
  rival_base,
  lag_availability,
  by = c("country", "season_start"),
  all.x = TRUE
)


# -------z score------
rival_vectors <- rival_base[
  ,
  {

    # ---- identify structural seasons ----
    min_season <- min(season_start)

    # ---- impute non-structural missing ----
    pts_t1_imp <- pts_t1
    pts_t2_imp <- pts_t2
    elo_imp    <- elo
    
    #q_t1  <- quantile(pts_t1, 0.25, na.rm = TRUE)
    #q_t2  <- quantile(pts_t2, 0.25, na.rm = TRUE)
    #q_elo <- quantile(elo,    0.25, na.rm = TRUE)

    elo_imp[is.na(elo_imp)] <- q_elo

    # ---- z-score scaling (after imputation) ----
    z_t  <- as.numeric(scale(pts_t))
    z_t1 <- as.numeric(scale(pts_t1_imp))
    z_t2 <- as.numeric(scale(pts_t2_imp))
    z_e  <- as.numeric(scale(elo_imp))

    # ---- set structural missing to 0 AFTER scaling ----
    if(season_start == 2017) {
      z_t1 <- 0
      z_t2 <- 0
    }
    
    if(season_start == 2018) {
      z_t2 <- 0
    }

    list(
      team_clean,
      z_pts_t  = z_t,
      z_pts_t1 = z_t1,
      z_pts_t2 = z_t2,
      z_elo    = z_e
    )
  },
  by = .(country, season_start)
]

```

```{r}
rival_vectors
```

```{r}
rival_base
```

### Rivalry score

Modified euclidean:

we only use the factors we have in the vector (pts_t, pts_t1, pts_t2, elo\> to calculate distance after z score normalisation. Arbitrarily assigning 0 to missing values messed up the rivalry score calculation.

```{r}
rivalry_scores <- rival_vectors[
  ,
  {
    X <- as.matrix(.SD)
    teams <- team_clean
    n <- nrow(X)

    out <- vector("list", n*(n-1)/2)
    idx <- 1L

    for (i in 1:(n-1)) {
      for (j in (i+1):n) {

        xi <- X[i, ]
        xj <- X[j, ]

        # identify observed dimensions
        # ignore dimension if BOTH are zero (structural zero)
        observed <- !(xi == 0 & xj == 0)

        k <- sum(observed)

        if (k == 0) {
          d <- NA_real_
        } else {
          d <- sqrt(sum((xi[observed] - xj[observed])^2) / k)
        }

        out[[idx]] <- list(
          team_1 = teams[i],
          team_2 = teams[j],
          rivalry_score = d
        )
        idx <- idx + 1L
      }
    }

    rbindlist(out)
  },
  by = .(country, season_start),
  .SDcols = c("z_pts_t", "z_pts_t1", "z_pts_t2", "z_elo")
]
rivalry_scores
```

Rolling rivalry score:

```{r}
rivalry_scores[
  ,
  `:=`(
    tA = pmin(team_1, team_2),
    tB = pmax(team_1, team_2)
  )
]

setorder(rivalry_scores, country, tA, tB, season_start)

alpha <- 0.25

rivalry_scores[
  ,
  rivalry_smooth :=
    Reduce(function(prev, curr) alpha * curr + (1 - alpha) * prev,
           rivalry_score,
           accumulate = TRUE),
  by = .(country, tA, tB)
]

rivalry_scores
```

### k closest

why knn?

1.  threshold - too generic; not even league or season specific
2.  quartile - lot of rivals missed; not smooth over seasons
3.  so we set knn - symmetric (only when both teams consider each other as rivals, we add them to the list)

```{r}
k <- 5

# 1) Expand to team-centric format
rival_long <- rbindlist(list(
  rivalry_scores[, .(country, season_start, team = tA, rival = tB, rivalry_smooth)],
  rivalry_scores[, .(country, season_start, team = tB, rival = tA, rivalry_smooth)]
))

# 2) k nearest per team
rival_knn <- rival_long[
  order(rivalry_smooth),
  head(.SD, k),
  by = .(country, season_start, team)
]

# 3) Keep only mutual rivals
rival_mutual <- rival_knn[
  rival_knn,
  on = .(country, season_start,
         team = rival,
         rival = team),
  nomatch = 0
]

# 4) Make undirected & unique
rival_undirected <- rival_mutual[
  ,
  .(
    country,
    season_start,
    team_a = pmin(team, rival),
    team_b = pmax(team, rival),
    is_rival = TRUE
  )
][
  ,
  .SD[1],
  by = .(country, season_start, team_a, team_b)
]
#rival_undirected[team_a == "barcelona" | team_b == "barcelona"]
fwrite(rival_undirected, out_rivals_csv)
```

winter transfer????

5610, 1042

## Transfers

We now get the transfer data for all teams in the top 5 leagues from 2018 to 2024 transfer windows.

```{r}
# Collect summer transfers (2018/19 .. 2024/25) using only country + season_start.
# Uses existing column `squad` in ~/sem3/research_project/rival_analysis/data/league_table.csv (no new "team" column created).

# ---------- PARAMETERS & PATHS ----------
league_csv <- "~/sem3/research_project/rival_analysis/data/league_table.csv"   # your uploaded file (use existing 'squad' col)
out_dir <- "~/sem3/research_project/rival_analysis/data"
cache_dir <- file.path(out_dir, "transfers_cache")
dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)

season_starts <- 2018:2024      # represent seasons 2018/19 .. 2024/25
sleep_sec <- 3                  # polite pause between requests
max_retries <- 0
retry_sleep <- 5

out_csv <- file.path(out_dir, "transfers_summer_2018_2025.csv")

# ---------- HELPERS ----------
safe_fn <- function(x) gsub("[^A-Za-z0-9_\\-]", "_", substr(x, 1, 120))

league_urls_cache_file <- function(country_name, start_year) {
  file.path(cache_dir, paste0("league_urls_", safe_fn(country_name), "_", start_year, ".rds"))
}
team_transfers_cache_file <- function(team_url, season_start) {
  file.path(cache_dir, paste0("team_transfers_", safe_fn(team_url), "_", season_start, ".rds"))
}

# cached call for league -> team urls
get_league_team_urls_cached <- function(country_name, start_year) {
  cachef <- league_urls_cache_file(country_name, start_year)
  if (file.exists(cachef)) return(readRDS(cachef))
  for (i in seq_len(max_retries)) {
    try({
      urls <- tm_league_team_urls(country_name = country_name, start_year = start_year)
      saveRDS(urls, cachef)
      Sys.sleep(sleep_sec)
      return(urls)
    }, silent = TRUE)
    message("Retrying league-team-urls for ", country_name, " ", start_year, " (attempt ", i, ")")
    Sys.sleep(retry_sleep)
  }
  warning("Failed to fetch league team URLs for ", country_name, " ", start_year)
  return(NULL)
}

# cached call for team -> transfers (summer)
get_team_transfers_cached <- function(team_url, season_start) {
  cachef <- team_transfers_cache_file(team_url, season_start)
  if (file.exists(cachef)) return(readRDS(cachef))
  for (i in seq_len(max_retries)) {
    try({
      xf <- tm_team_transfers(team_url = team_url, transfer_window = "summer")
      saveRDS(xf, cachef)
      Sys.sleep(sleep_sec)
      return(xf)
    }, silent = TRUE)
    message("Retrying team transfers for ", team_url, " (attempt ", i, ")")
    Sys.sleep(retry_sleep)
  }
  message("No transfers found for team: ", team_url)
  return(NULL)
}

# cheap fuzzy compare to decide whether a returned team name matches a squad in your league table
close_name_match <- function(x, candidates, max_dist = 0.18) {
  # x: single string, candidates: vector of strings
  if (is.na(x) || x == "") return(FALSE)
  # use jaro-winkler via stringdist if available; fallback to grepl substring match
  if (requireNamespace("stringdist", quietly = TRUE)) {
    d <- stringdist::stringdist(tolower(str_squish(x)), tolower(str_squish(candidates)), method = "jw")
    any(d <= max_dist)
  } else {
    any(grepl(tolower(str_squish(x)), tolower(candidates), fixed = TRUE))
  }
}

# ---------- 1) Read league table and derive country+season_start combos ----------
dt <- fread(league_csv)
# use existing 'squad' column (do not create new 'team' column)
if (!"squad" %in% names(dt)) stop("'squad' column not found in league_table.csv")
dt[, squad_clean := tolower(str_squish(squad))]
dt[, season_start := as.integer(sub("([0-9]{4}).*", "\\1", season))]   # extracts starting year like 2018 from "2018/19"
dt <- dt[season_start %in% season_starts]

country_seasons <- unique(dt[, .(country, season_start)])
setorder(country_seasons, country, season_start)

# Build quick lookup of squads per country+season for filtering
squads_by_country_season <- dt[, .(squad_list = list(unique(squad_clean))), by = .(country, season_start)]
setkey(squads_by_country_season, country, season_start)

# ---------- 2) iterate country+season, fetch team URLs, then fetch summer transfers ----------
all_transfers <- list()
for (r in seq_len(nrow(country_seasons))) {
  cs <- country_seasons[r]
  cntry <- cs$country
  sstart <- cs$season_start
  message("Processing: ", cntry, " | season_start ", sstart)
  urls_tbl <- get_league_team_urls_cached(cntry, sstart)
  if (is.null(urls_tbl) || length(urls_tbl) == 0) {
    message(" No team URLs returned for country ", cntry, " season ", sstart, " — skipping")
    next
  }

  # normalize returned structure to a data.table with at least a team_url column and optionally a display name
  if (is.character(urls_tbl)) {
    team_urls_df <- unique(data.table(team_url = urls_tbl))
  } else if (is.data.frame(urls_tbl) || is.data.table(urls_tbl)) {
    # try to detect url column and name column
    url_col <- intersect(names(urls_tbl), c("team_url", "url", "link", "teamlink", "team_url_href"))[1]
    name_col <- intersect(names(urls_tbl), c("team", "club", "club_name", "team_name", "name"))[1]
    if (!is.na(url_col)) {
      team_urls_df <- unique(data.table(team_url = as.character(urls_tbl[[url_col]])))
      if (!is.na(name_col)) team_urls_df[, team_name := as.character(urls_tbl[[name_col]])]
    } else {
      # fallback: use first column as url
      team_urls_df <- unique(data.table(team_url = as.character(urls_tbl[[1]])))
    }
  } else {
    message(" Unrecognized return type from tm_league_team_urls() — skipping this country-season.")
    next
  }

  # OPTIONAL: filter returned URLs to squads that exist in your league_table for that country+season
  # This avoids scraping unrelated clubs. Keep enabled by default.
  filter_local <- TRUE
  if (filter_local && "team_name" %in% names(team_urls_df)) {
    squads_local <- squads_by_country_season[J(cntry, sstart), squad_list][[1]]
    # keep only urls whose team_name roughly matches any squad
    keep_mask <- sapply(team_urls_df$team_name, function(nm) close_name_match(nm, squads_local))
    if (all(!keep_mask)) {
      # if filter removed all, fall back to not filtering (safer)
      message("  Local filtering removed all returned team URLs — falling back to unfiltered list")
      team_urls_final <- unique(team_urls_df$team_url)
    } else {
      team_urls_final <- unique(team_urls_df$team_url[keep_mask])
    }
  } else {
    # if no display name provided by the function, use all returned urls
    team_urls_final <- unique(team_urls_df$team_url)
  }

  # for each team_url, fetch summer transfers (cached)
  for (tu in team_urls_final) {
    message("  Fetching summer transfers for: ", tu)
    xf <- get_team_transfers_cached(tu, sstart)
    if (is.null(xf)) next
    dt_xf <- as.data.table(xf)
    # add context columns so you know which country+season they belong to
    dt_xf[, team_url := tu]
    dt_xf[, country := cntry]
    dt_xf[, season_start := sstart]
    setnames(dt_xf, tolower(names(dt_xf)))
    all_transfers[[length(all_transfers) + 1]] <- dt_xf
  }
}

# ---------- 3) combine and save ----------
if (length(all_transfers) == 0) {
  warning("No transfers fetched. Check that tm_league_team_urls() and tm_team_transfers() work in your environment.")
} else {
  transfers_dt <- rbindlist(all_transfers, use.names = TRUE, fill = TRUE)
  # add season numeric and save
  transfers_dt[, season := season_start]
  fwrite(transfers_dt, out_csv)
  message("Saved summer transfers to: ", out_csv)
  message("Cached per-team rds files are in: ", cache_dir)
}

```

### Filtering out transfers - only within teams from the top5 leagues

The above data contains transfers that happened for teams from the top 5 leagues. But, it doesn't necessarily mean that all such transfers happened within the top 5 leagues. Teams also buy players from other leagues. So, we add a filter such that both the seller and buyer teams must be within the top 5 leagues for our observation.

```{r}
transfers_path <- "~/sem3/research_project/rival_analysis/data/transfers_summer_2018_2025.csv"
transfers  <- fread(transfers_path)
league_table <- fread("~/sem3/research_project/rival_analysis/data/league_table.csv")
top5_leagues <- c("Premier League", "LaLiga", "Bundesliga", "Ligue 1", "Serie A")

top5_league_transfers <- filter(transfers,
       league %in% top5_leagues,
       league_2 %in% top5_leagues,
       is_loan == FALSE,
       transfer_type == "Arrivals")
fwrite(top5_league_transfers, "~/sem3/research_project/rival_analysis/data/top5_league_transfers.csv")
```

## Mapping

Before we proceed with analysis, we try to map the team names between the different tables we have gathered so far.

```{r}

## ===================== PATHS ===================== ##
rivals_path    <- "~/sem3/research_project/rival_analysis/data/rivalry_lookup.csv"
transfers_path <- "~/sem3/research_project/rival_analysis/data/top5_league_transfers.csv"
league_path <- "~/sem3/research_project/rival_analysis/data/league_table.csv"
out_dir     <- "~/sem3/research_project/rival_analysis/data"

## ===================== 1. LOAD DATA ===================== ##
league <- fread(league_path)
xf     <- fread(transfers_path)

## ===================== 2. BUILD CANONICAL TEAM LIST ===================== ##
# League table defines truth
league[, team_clean := tolower(str_squish(squad))]
league[, country := as.character(country)]

canon_teams <- unique(
  league[, .(country, team_clean)]
)

## ===================== 3. EXTRACT RAW TRANSFER TEAM NAMES ===================== ##
xf[, buyer_raw  := tolower(str_squish(team_name))]
xf[, seller_raw := tolower(str_squish(club_2))]
xf[, country    := as.character(country)]

raw_teams <- unique(rbind(
  xf[, .(country, raw = buyer_raw)]
))

valid_canon <- canon_teams[, .(country, team_clean)]

raw_teams <- raw_teams[
  country %in% valid_canon$country
]

normalize <- function(x) {
  x |>
    tolower() |>
    str_replace_all("&", "and") |>
    str_replace_all("[^a-z ]", "") |>
    str_squish()
}

valid_canon[, team_norm := normalize(team_clean)]
raw_teams[, raw_norm := normalize(raw)]


## ===================== 4. FUZZY MATCH (RAW → CANONICAL) ===================== ##
team_map <- stringdist_inner_join(
  raw_teams,
  valid_canon,
  by = c("raw" = "team_clean"),
  method = "jw",
  max_dist = 0.2,
  distance_col = "dist"
)

# IMPORTANT: force data.table
setDT(team_map)

## ===================== 5. FIX country.x / country.y ===================== ##
# Keep only same-country matches
team_map <- team_map[country.x == country.y]

# Clean column names
setnames(team_map, "country.x", "country")
team_map[, country.y := NULL]
team_map[, country.dist := NULL]
team_map[, raw.dist := NULL]

# Keep best match per (country, raw)
setorder(team_map, country, raw, dist)
team_map <- team_map[, .SD[1], by = .(country, raw)]


## ===================== 6. QA CHECKS ===================== ##
# Unmatched raw names (inspect & manually fix if needed)
unmatched <- raw_teams[!raw %in% team_map$raw]
# fwrite(unmatched, file.path(out_dir, "unmatched_transfer_teams.csv"))

cat("Unmatched raw team names saved to unmatched_transfer_teams.csv\n")
```

```{r}
## ===================== 7. APPLY MAPPING TO TRANSFERS ===================== ##
manual_map <- fread("~/sem3/research_project/rival_analysis/data/unmatched_transfer_teams.csv")

# Rivalries: teams
riv <- fread(rivals_path)
riv[, team_a_clean := tolower(str_squish(team_a))]
riv[, team_b_clean := tolower(str_squish(team_b))]
riv[, season       := as.integer(season_start)]

## ===================== 3. BUILD DEDUPLICATED RIVAL LOOKUP ===================== ##
# Collapse to unique rival pairs per country-season-teamA-teamB
riv_unique <- riv[
  is_rival == TRUE,
  .(country, season, team_a_clean, team_b_clean)
]

# Ensure uniqueness (this removes pos+elo duplicates)
riv_unique <- unique(riv_unique)

# Build bidirectional lookup
riv_long <- rbind(
  riv_unique[, .(country, season, from = team_a_clean, to = team_b_clean, is_rival = TRUE)],
  riv_unique[, .(country, season, from = team_b_clean, to = team_a_clean, is_rival = TRUE)]
)

team_map_final <- unique(rbind(
  team_map[, .(country, raw, team_clean)],
  manual_map
))

xf <- merge(
  xf,
  team_map_final,
  by.x = c("country", "buyer_raw"),
  by.y = c("country", "raw"),
  all.x = TRUE
)
setnames(xf, "team_clean", "buyer_clean")
dup_last <- duplicated(names(xf), fromLast = TRUE)
xf <- xf[, !dup_last, with = FALSE]
# Sellers (needed for rivalry tagging)
# xf <- merge(
#  xf,
#  team_map_final,
# by.x = c("country", "seller_raw"),
# by.y = c("country", "raw"),
#  all.x = TRUE
#)
setnames(xf, "seller_raw", "seller_clean")
dup <- duplicated(names(xf))
xf <- xf[, !dup, with = FALSE]

xf[, is_rival := NULL]
xf <- merge(
  xf,
  riv_long,
  by.x = c("country", "season", "seller_clean", "buyer_clean"),
  by.y = c("country", "season", "from", "to"),
  all.x = TRUE
)
xf[, is_rival := !is.na(is_rival)]
xf[, pos_clean        := tolower(str_squish(player_position))]

fwrite(xf, "~/sem3/research_project/rival_analysis/data/top5_league_transfers_clean.csv")
```

## Opportunity Set

Next, we build the opportunity set. We try to find alternative choices for a rival transfer. We count:

-   Total rival options available (rivalry is local to a league)

-   Non-rival options available within the league

-   Global non-rival options (top 5 leagues)

```{r}

# Load data
transfers <- read_csv("~/sem3/research_project/rival_analysis/data/top5_league_transfers_clean.csv")
rivalry_lookup <- read_csv("~/sem3/research_project/rival_analysis/data/rivalry_lookup.csv")
transfers <- transfers %>%
  mutate(transfer_id = row_number())

# -------------------------
# Helper function to get rivals of a team
# -------------------------
get_rivals <- function(team_name, rivalry_df) {
  rivalry_df %>%
    filter(team_a == team_name | team_b == team_name) %>%
    mutate(rival = if_else(team_a == team_name, team_b, team_a)) %>%
    pull(rival) %>%
    unique()
}

# -------------------------
# Only evaluate rival transfers
# -------------------------
rival_transfers <- transfers %>%
  filter(is_rival == TRUE)

# -------------------------
# Main loop (robust version)
# -------------------------
results <- rival_transfers %>%
  rowwise() %>%
  mutate(

    # Get rival clubs of buyer
    rivals_vec = list(get_rivals(buyer_clean, rivalry_lookup)),

    # -----------------------
    # League-constrained pool
    # -----------------------
    league_pool = list(
      transfers %>%
        filter(
          transfer_id != transfer_id,
          season == season,
          pos_clean == pos_clean,
          country_2 == country
        )
    ),

    # Count rival options
    total_rival_options = sum(
      transfers$transfer_id != transfer_id &
      transfers$season == season &
      transfers$pos_clean == pos_clean &
      transfers$country_2 == country &
      transfers$seller_clean %in% rivals_vec
    ),

    # Count non-rival options
    non_rival_options = sum(
      transfers$transfer_id != transfer_id &
      transfers$season == season &
      transfers$pos_clean == pos_clean &
      transfers$country_2 == country &
      !(transfers$seller_clean %in% rivals_vec)
    ),

    # -----------------------
    # Global pool (no league + no rival restriction)
    # -----------------------
    global_pool_options = sum(
      transfers$transfer_id != transfer_id &
      transfers$season == season &
      transfers$pos_clean == pos_clean
    )

  ) %>%
  ungroup() %>%
  select(-league_pool)
```

```{r}
options_pool <- results %>%
  select(
    country,
    season,
    buyer_clean,
    seller_clean,
    player_name,
    pos_clean,
    non_rival_options,      # league restricted
    global_pool_options,    # league-agnostic
    total_rival_options
  )
fwrite(options_pool, "~/sem3/research_project/rival_analysis/data/options_pool.csv")
```

```{r}
options_pool
```

```{r}
#filter(top5_league_transfers,
#       league_2 == "Premier League",
#       pos_clean == "centre-back",
#       season == 2019)
```

## Conditional Multinomial choice model

Within each transfer decision, are rival sellers more likely to be the chosen alternative, controlling for player characteristics?

```{r}
transfers <- read_csv("~/sem3/research_project/rival_analysis/data/top5_league_transfers_clean.csv")

transfers <- transfers %>%
  mutate(transfer_id = row_number())

choice_data <- transfers %>%
  #filter(is_rival == TRUE) %>%
  rowwise() %>%
  do({

    current <- .
    
    pool <- transfers %>%
      filter(
        season == current$season,
        pos_clean == current$pos_clean,
        #country_2 == current$country,
        transfer_id != current$transfer_id
      )
    
    # Add chosen player back into pool
    pool <- bind_rows(pool, current)
    
    pool %>%
      mutate(
        choice_id = current$transfer_id,
        chosen = ifelse(transfer_id == current$transfer_id, 1, 0)
      )
  }) %>% 
  ungroup()

choice_data <- choice_data %>%
  mutate(
    nat_match = ifelse(player_nationality == country, 1, 0),
    log_fee = log(transfer_fee + 1)
  )

get_rivals <- function(team_name, rivalry_df) {
  rivalry_df %>%
    filter(team_a == team_name | team_b == team_name) %>%
    mutate(rival = if_else(team_a == team_name, team_b, team_a)) %>%
    pull(rival) %>%
    unique()
}

choice_data <- choice_data %>%
  rowwise() %>%
  mutate(
    rival = ifelse(
      seller_clean %in% get_rivals(buyer_clean, rivalry_lookup),
      1, 0
    )
  ) %>%
  ungroup()
```

```{r}
model <- clogit(
  rival ~ player_age +
    minutes_played + goals + nat_match +
    log_fee + 
    chosen +
    #rival + 
    strata(choice_id),
  data = choice_data
)

summary(model)
```

```{r}
fwrite(choice_data,"~/sem3/research_project/rival_analysis/data/choice_data.csv")
choice_data
```

Within each transfer decision (choice set), how do player characteristics and rivalry affect the probability that a player is chosen?

```{r}
model2 <- clogit(
  chosen ~ player_age +
    minutes_played + goals + nat_match +
    log_fee + 
    rival + 
    strata(choice_id),
  data = choice_data
)

summary(model2)
```
