---
title: "What motivates a transfer: player performance, transfer fee or rivalry?"
author: "Vignesh Thiruvel Selvam"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(worldfootballR)
library(future)
library(furrr)
library(dplyr)
library(data.table)
library(stringr)
library(httr)
library(readr)
library(fuzzyjoin)
library(geosphere)
library(survival)
#library(ggplot2)
```

## Getting League data for each season

We get league table data for the top 5 leagues from 2017/18 season to 2024/2025 season.

```{r}
league_path <- "data/league_table.csv"
if (!file.exists(league_path)) {
  leagues <- c("England", "Spain", "Germany", "Italy", "France")
  seasons <- 2017:2024
  league_season_combos <- expand.grid(league = leagues, season = seasons)
  
  # Parallel fetching
  results <- future_pmap_dfr(league_season_combos, function(league, season) {
    tryCatch({
      matchday <- case_when(
        league == "Germany" ~ 34,
        league == "France" & season == 2019 ~ 27,
        league == "France" & season >= 2023 ~ 34,
        TRUE ~ 38
      )
      message(paste("Fetching:", league, season))
      df <- tm_matchday_table(country_name=league, start_year=season, matchday=matchday)
      df$season <- season
      df
    }, error = function(e) {
      message(paste("Error fetching", league, season, ":", e$message))
      return(NULL)
    })
  })
  results[, league_name := league |> str_remove("Games of Matchday \\d+ ") |> str_remove(" \\d{2}/\\d{2}")]
  fwrite(results, league_path)
}
```

## Rivalry

Rivalry is defined based on the followin strategy:

1.  We already have the league table history for teams. We now get the ELO rating of teams for all these seasons.
2.  Next, we define a vector: \<points scored by a team this season, points scored by the team previous season, points scored by the team 2 seasons ago, current ELO\>.
3.  Based on this, we calculate distances between teams. Note: We ignore missing data altogether. We have 2 such cases:
    1.  We only have the data starting from the end of the 2017/18 season. Thus, for this season, the 2nd and 3rd values in the vector are not present. We leave this as is.
    2.  Also, we could have missing data for a team that was newly promoted to the league. We again ignore missing values in the vector.
4.  Next, we attribute this distance score over multiple seasons to smoothen out the rivalry score, so that there aren't drastic changes in rivalry across seasons.
5.  Next, we calculate a continuous rivalry proximity score, a rank-based percentile of the smoothened distance. Then, we invert it so that when the score is closer to 1, then the 2 teams are closer rivals.

```{r}

# 1) read
dt_raw <- fread("data/league_table.csv")

dt_raw[, team := str_squish(squad)]
dt_raw[, season := as.character(season)]
dt_raw[, league := as.character(league)]
dt_raw[, pos := as.integer(rk)]

if ("matchday" %in% names(dt_raw)) {
  dt_final <- dt_raw[!is.na(pos), .SD[which.max(as.numeric(matchday))], 
                     by = .(country, league, season, team)]
} else {
  dt_final <- dt_raw[!is.na(pos), .SD[1], by = .(country, league, season, team)]
}
# Keep only needed cols
dt_final <- dt_final[, .(country, league, season, team, pos, pts)]
dt_final[, team_low := tolower(team)]
dt_final[, country_low := tolower(country)]
dt_final[, league_low := tolower(league)]
```

### Getting ELO

```{r}
# ---------------- PARAMETERS ----------------
seasons <- 2018:2025
snapshot_month_day <- "-09-01"
max_name_dist <- 0.18

# ---------------- PATHS (uses uploaded files in data) ----------------
league_csv <- "data/league_table.csv"
manual_map_path <- "data/manual_club_map.csv"
out_dir <- "data"
out_rivals_csv <- file.path(out_dir, "rivalry_lookup.csv")
out_elo_raw_dir <- file.path(out_dir, "elo_snapshots")
dir.create(out_elo_raw_dir, showWarnings = FALSE, recursive = TRUE)

# ---------------- 1) load & prepare league table (dt_final) ----------------
dt_raw <- fread(league_csv)
dt_raw[, team := str_squish(squad)]
dt_raw[, country := as.character(country)]
dt_raw[, league := as.character(league)]
dt_raw[, season := as.character(season)]
dt_raw[, pos := as.integer(rk)]

if ("matchday" %in% names(dt_raw)) {
  dt_final <- dt_raw[!is.na(pos), .SD[which.max(as.numeric(matchday))], by = .(country, league, season, team)]
} else {
  dt_final <- dt_raw[!is.na(pos), .SD[1], by = .(country, league, season, team)]
}
dt_final <- dt_final[, .(country, league, season, team, pos)]
dt_final[, team_clean := tolower(str_squish(team))]
# Precompute season_start string on dt_final for lookups (e.g., "2017" from "2017/18")
dt_final[, season_start_str := sub("([0-9]{4}).*", "\\1", season)]

# ---------------- 2) fetch ClubElo snapshots (cached) ----------------
fetch_elo_for_date <- function(date_str) {
  destfile <- file.path(out_elo_raw_dir, paste0("clubelo_", date_str, ".csv"))
  if (file.exists(destfile)) {
    message("Using cached Elo snapshot: ", destfile)
    df <- read_csv(destfile, show_col_types = FALSE)
    setDT(df); return(df)
  }
  url <- paste0("http://api.clubelo.com/", date_str)
  message("Fetching from API: ", url)
  r <- httr::GET(url)
  if (httr::status_code(r) != 200) stop("ClubElo request failed for date ", date_str, " (status ", httr::status_code(r), ")")
  txt <- httr::content(r, as = "text", encoding = "UTF-8")
  df <- read_csv(txt, show_col_types = FALSE)
  setDT(df)
  setnames(df, tolower(names(df)))
  club_col <- grep("club|team|name", names(df), value = TRUE)[1]
  elo_col  <- grep("elo|rating", names(df), value = TRUE)[1]
  if (is.na(club_col) || is.na(elo_col)) stop("Unable to find club/elo columns in ClubElo CSV for date: ", date_str)
  df <- df[, .(club = get(club_col), elo = as.numeric(get(elo_col)))]
  df <- df[!is.na(elo) & !is.na(club)]
  df[, club_clean := tolower(str_squish(gsub("\\(.*\\)", "", club)))]
  df <- df[order(-elo)]
  df[, snapshot_date := date_str]
  write_csv(df, destfile)
  message("Saved snapshot: ", destfile)
  return(df)
}

# Fetch and cache all snapshots (only does network calls if missing)
elo_all <- rbindlist(lapply(seasons, function(y) {
  d <- paste0(y, snapshot_month_day)
  message("Fetch snapshot: ", d)
  Sys.sleep(0.5) # be kind to API
  fetch_elo_for_date(d)
}), use.names = TRUE, fill = TRUE)
message("Total Elo rows fetched (or cached): ", nrow(elo_all))
elo_all[, season_start := as.character(as.integer(substr(snapshot_date, 1, 4)) - 1L)]

# ---------------- 3) APPLY MANUAL MAPPINGS (rowwise, robust) ----------------
manual_map_applied <- data.table()  # result holder
if (file.exists(manual_map_path)) {
  mm <- fread(manual_map_path)
  # normalize columns: support snapshot_date or season_start
  if (!("snapshot_date" %in% names(mm)) & ("season_start" %in% names(mm))) {
    mm[, snapshot_date := paste0(season_start, snapshot_month_day)]
  }
  if (!("club" %in% names(mm)) | !("mapped_team" %in% names(mm))) {
    stop("manual_club_map.csv must contain at minimum: club, mapped_team, and snapshot_date or season_start")
  }
  mm[, club_clean := tolower(str_squish(gsub("\\(.*\\)", "", club)))]
  mm[, mapped_team_clean := tolower(str_squish(mapped_team))]
  mm[, season_start := as.character(substring(snapshot_date, 1, 4))]
  # rowwise validate and build manual_map_applied
  valid_rows <- list()
  for (i in seq_len(nrow(mm))) {
    row <- mm[i]
    ss <- as.character(row$season_start)
    mt_clean <- row$mapped_team_clean
    # check dt_final: does mapped team exist in that season_start?
    cand <- dt_final[team_clean == mt_clean & season_start_str == ss]
    if (nrow(cand) == 0L) next  # invalid mapping: mapped_team not present in league for that season
    # find matching elo rows for the same snapshot + club_clean
    erows <- elo_all[season_start == ss & club_clean == row$club_clean]
    if (nrow(erows) == 0L) next  # no matching elo row to attach
    # build manual match rows
    for (er in seq_len(nrow(erows))) {
      for (cd in seq_len(nrow(cand))) {
        valid_rows[[length(valid_rows) + 1]] <- data.table(
          snapshot_date = erows[er, snapshot_date],
          club = erows[er, club],
          club_clean = erows[er, club_clean],
          elo = erows[er, elo],
          team = cand[cd, team],
          team_clean = cand[cd, team_clean],
          country = cand[cd, country],
          league = cand[cd, league],
          dist = 0  # mark manual mapping as exact
        )
      }
    }
  }
  if (length(valid_rows) > 0) {
    manual_map_applied <- rbindlist(valid_rows, use.names = TRUE, fill = TRUE)
    message("Applied ", nrow(manual_map_applied), " manual mapping rows from: ", manual_map_path)
  } else {
    message("No valid manual mappings found in: ", manual_map_path)
  }
} else {
  message("No manual_club_map.csv found at: ", manual_map_path, " (continuing without manual mappings)")
}

# ---------------- 4) FUZZY MATCH remaining Elo clubs -> dt_final teams ----------------
# Build candidate list excluding manual-mapped (by snapshot_date+club_clean)
elo_to_match <- copy(elo_all[, .(snapshot_date, club, club_clean, elo, season_start)])
if (nrow(manual_map_applied) > 0) {
  keys_mm <- unique(manual_map_applied[, .(snapshot_date, club_clean)])
  setkey(elo_to_match, snapshot_date, club_clean)
  setkey(keys_mm, snapshot_date, club_clean)

  # Proper anti-join (no need for fsetdiff)
  elo_to_match <- elo_to_match[!keys_mm]
}

# Build teams by season_start
teams_by_start <- unique(dt_final[, .(season_start_str, team, team_clean, country, league)])
setnames(teams_by_start, "season_start_str", "season_start")

matched_list <- list()
unmatched_list <- list()
for (yr in sort(unique(elo_to_match$season_start))) {
  eloy <- elo_to_match[season_start == yr, .(snapshot_date, club, club_clean, elo)]
  targ <- teams_by_start[season_start == yr, .(team, team_clean, country, league)]
  if (nrow(targ) == 0L) { unmatched_list[[yr]] <- eloy; next }
  mj <- stringdist_inner_join(eloy, targ, by = c("club_clean" = "team_clean"),
                              method = "jw", max_dist = max_name_dist, distance_col = "dist")
  setDT(mj); setorder(mj, snapshot_date, club, dist)
  if (nrow(mj) > 0) {
    mj <- mj[, .SD[1], by = .(snapshot_date, club)]  # best match per elo club
    matched_list[[yr]] <- mj
    matched_clubs <- unique(mj$club)
    unmatched_list[[yr]] <- eloy[!club %in% matched_clubs]
  } else {
    unmatched_list[[yr]] <- eloy
  }
}

matched_dt_fuzzy <- if (length(matched_list) > 0) rbindlist(matched_list, use.names = TRUE, fill = TRUE) else data.table()
unmatched_dt <- if (length(unmatched_list) > 0) rbindlist(unmatched_list, use.names = TRUE, fill = TRUE) else data.table()

# Combine manual-applied matches and fuzzy matches into matched_dt
if (nrow(manual_map_applied) > 0 && nrow(matched_dt_fuzzy) > 0) {
  matched_dt <- rbindlist(list(manual_map_applied, matched_dt_fuzzy[, .(snapshot_date, club, club_clean, elo, team, team_clean, country, league, dist)]),
                          use.names = TRUE, fill = TRUE)
} else if (nrow(manual_map_applied) > 0) {
  matched_dt <- copy(manual_map_applied)
} else {
  matched_dt <- copy(matched_dt_fuzzy)
}

# Save remaining unmatched for inspection
if (nrow(unmatched_dt) > 0) {
  fwrite(unique(unmatched_dt), file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
  message("Wrote remaining unmatched Elo clubs to: ", file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
} else {
  # produce empty file to indicate nothing remaining
  fwrite(data.table(snapshot_date=character(), club=character(), club_clean=character(), elo=numeric()),
         file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
  message("No remaining unmatched Elo clubs; created empty file: ", file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
}

# ---------------- 5) set Elo-season = snapshot_year - 1 (numeric) for Elo outputs ONLY ----------------
if (nrow(matched_dt) == 0) {
  message("Warning: matched_dt is empty (no Elo->league matches). Elo-only will be empty.")
}
matched_dt[, snapshot_year := as.integer(substr(snapshot_date, 1, 4))]
matched_dt[, elo_season := snapshot_year - 1L]  # e.g., 2018 -> 2017
```

### Creating Rivalry Vector

```{r}

# ------Prepare points history-----
# Extract points table
pts_dt <- copy(dt_raw)

# normalize columns
pts_dt[, team_clean := tolower(str_squish(squad))]
pts_dt[, pts := as.numeric(pts)]

# keep only valid rows
pts_dt <- pts_dt[!is.na(pts),
                 .(country, league, season_start = season, team_clean, pts)]
pts_dt[, season_start := as.integer(season_start)]

# -------Build lagged points-------
# current season points
pts_t <- pts_dt[,.(season_start, team_clean, pts)]
setnames(pts_t, "pts", "pts_t")

# t-1
pts_t1 <- pts_dt[,.(season_start, team_clean, pts)]
pts_t1[, season_start := season_start + 1]
setnames(pts_t1, "pts", "pts_t1")

# t-2
pts_t2 <- pts_dt[,.(season_start, team_clean, pts)]
pts_t2[, season_start := season_start + 2]
setnames(pts_t2, "pts", "pts_t2")


# -----assemble base rivalry------

rival_base <- unique(
  pts_dt[, .(country, league, season_start, team_clean)]
)

# merge points
rival_base <- merge(rival_base, pts_t,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- merge(rival_base, pts_t1,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- merge(rival_base, pts_t2,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)


# ------attach elo-------
elo_dt <- matched_dt[
  , .(season_start = elo_season,
      team_clean, elo)
]

rival_base <- merge(rival_base, elo_dt,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- unique(rival_base, by = c("team_clean", "season_start"))

elo_quartiles <- rival_base[
  , .(q_elo = quantile(elo, 0.25, na.rm = TRUE)),
  by = .(season_start)
]

rival_base <- merge(rival_base, elo_quartiles,
                    by = "season_start",
                    all.x = TRUE)

# rival_base[is.na(elo), elo := q_elo]


# flag seasons where lagged points are structurally missing
lag_availability <- rival_base[
  , .(
      has_pts_t1 = any(!is.na(pts_t1)),
      has_pts_t2 = any(!is.na(pts_t2))
    ),
  by = .(country, season_start)
]

rival_base <- merge(
  rival_base,
  lag_availability,
  by = c("country", "season_start"),
  all.x = TRUE
)


# -------z score------
rival_vectors <- rival_base[
  ,
  {

    # ---- identify structural seasons ----
    min_season <- min(season_start)

    # ---- impute non-structural missing ----
    pts_t1_imp <- pts_t1
    pts_t2_imp <- pts_t2

    # ---- z-score scaling (after imputation) ----
    z_t  <- as.numeric(scale(pts_t))
    z_t1 <- as.numeric(scale(pts_t1_imp))
    z_t2 <- as.numeric(scale(pts_t2_imp))
    z_e  <- as.numeric(scale(elo))

    # ---- set structural missing to 0 AFTER scaling ----
    if(season_start == 2017) {
      z_t1 <- NA_real_
      z_t2 <- NA_real_
    }
    
    if(season_start == 2018) {
      z_t2 <- NA_real_
    }

    list(
      team_clean,
      z_pts_t  = z_t,
      z_pts_t1 = z_t1,
      z_pts_t2 = z_t2,
      z_elo    = z_e
    )
  },
  by = .(country, season_start)
]

```

```{r}
rival_vectors
```

```{r}
rival_base
```

### Calculating Rivalry score

Modified euclidean:

we only use the factors we have in the vector (pts_t, pts_t1, pts_t2, elo\> to calculate distance after z score normalisation. Arbitrarily assigning 0 to missing values messed up the rivalry score calculation.

```{r}
rivalry_scores <- rival_vectors[
  ,
  {
    X <- as.matrix(.SD)
    teams <- team_clean
    n <- nrow(X)

    out <- vector("list", n*(n-1)/2)
    idx <- 1L

    for (i in 1:(n-1)) {
      for (j in (i+1):n) {

        xi <- X[i, ]
        xj <- X[j, ]

        # identify observed dimensions
        # ignore dimension if BOTH are NA (structurally missing)
        observed <- !(is.na(xi) & is.na(xj))

        k <- sum(observed)

        if (k == 0) {
          d <- NA_real_
        } else {
          d <- sqrt(sum((xi[observed] - xj[observed])^2) / k)
        }

        out[[idx]] <- list(
          team_1 = teams[i],
          team_2 = teams[j],
          rivalry_score = d
        )
        idx <- idx + 1L
      }
    }

    rbindlist(out)
  },
  by = .(country, season_start),
  .SDcols = c("z_pts_t", "z_pts_t1", "z_pts_t2", "z_elo")
]
rivalry_scores
```

Smoothened rivalry score:

```{r}
rivalry_scores[
  ,
  `:=`(
    tA = pmin(team_1, team_2),
    tB = pmax(team_1, team_2)
  )
]

setorder(rivalry_scores, country, tA, tB, season_start)

alpha <- 0.25

rivalry_scores[
  ,
  rivalry_smooth :=
    Reduce(function(prev, curr) alpha * curr + (1 - alpha) * prev,
           rivalry_score,
           accumulate = TRUE),
  by = .(country, tA, tB)
]

rivalry_scores
```

### Continuous Rivalry Proximity

```{r}
# Expand to team-centric format (same as before)
rival_long <- rbindlist(list(
  rivalry_scores[, .(country, season_start, team = tA, rival = tB, rivalry_smooth)],
  rivalry_scores[, .(country, season_start, team = tB, rival = tA, rivalry_smooth)]
))

rival_sym <- rival_long[
  rival_long,
  on = .(country, season_start,
         team = rival,
         rival = team),
  nomatch = 0,
  allow.cartesian = TRUE
][
  ,
  .(
    country,
    season_start,
    team_a = pmin(team, rival),
    team_b = pmax(team, rival),
    rivalry_symmetric = mean(c(rivalry_smooth, i.rivalry_smooth))
  ),
  by = .(country, season_start, team, rival)
][
  ,
  .SD[1],
  by = .(country, season_start, team_a, team_b)
]

rival_sym[
  ,
  rivalry_percentile := frank(rivalry_symmetric, ties.method = "average") / .N,
  by = .(country, season_start)
]

# Invert so 1 = closest rival
rival_sym[
  ,
  rivalry_proximity := 1 - rivalry_percentile,
  by = .(country, season_start)
]

rival_continuous <- rival_sym[
  ,
  .(country, season_start, team_a, team_b, rivalry_proximity)
]

fwrite(rival_continuous, "data/rivalry_continuous.csv")
rival_continuous[order(rivalry_proximity, decreasing = TRUE)]
```

### k closest

\*(right now knn is calculated and not used anywhere)

why knn?

1.  threshold - too generic; not even league or season specific
2.  quartile - lot of rivals missed; not smooth over seasons
3.  so we set knn - symmetric (only when both teams consider each other as rivals, we add them to the list)

```{r}
k <- 5

# 1) Expand to team-centric format
rival_long <- rbindlist(list(
  rivalry_scores[, .(country, season_start, team = tA, rival = tB, rivalry_smooth)],
  rivalry_scores[, .(country, season_start, team = tB, rival = tA, rivalry_smooth)]
))

# 2) k nearest per team
rival_knn <- rival_long[
  order(rivalry_smooth),
  head(.SD, k),
  by = .(country, season_start, team)
]

# 3) Keep only mutual rivals
rival_mutual <- rival_knn[
  rival_knn,
  on = .(country, season_start,
         team = rival,
         rival = team),
  nomatch = 0
]

# 4) Make undirected & unique
rival_undirected <- rival_mutual[
  ,
  .(
    country,
    season_start,
    team_a = pmin(team, rival),
    team_b = pmax(team, rival),
    is_rival = TRUE
  )
][
  ,
  .SD[1],
  by = .(country, season_start, team_a, team_b)
]
#rival_undirected[team_a == "barcelona" | team_b == "barcelona"]
fwrite(rival_undirected, out_rivals_csv)
```

## Transfers

We now get the transfer data for all teams in the top 5 leagues from 2018 to 2024 transfer windows. We only fetch the summer window transfers. We don't consider winter transfers for our models because:

1.  Winter transfers are based on a team's position in the middle of a season. Often, it is considered to be a temporary solution (to fill up an injured player's spot or a desperate transfer to avoid relegation, etc.,.).
2.  It could even affect a team's options in the next summer window, due to constrained budget (already spent on players in the winter transfer), overcrowded positions (a player was bought due to an injury to an existing player, but the injured player has now returned, etc.).

Thus, it is altogether a different strategy compared to summer transfers and beyond the scope of what we are trying to establish in this analysis.

```{r}
# Collect summer transfers (2018/19 .. 2024/25) using only country + season_start.
# Uses existing column `squad` in data/league_table.csv (no new "team" column created).

# ---------- PARAMETERS & PATHS ----------
league_csv <- "data/league_table.csv"   # your uploaded file (use existing 'squad' col)
out_dir <- "data"
cache_dir <- file.path(out_dir, "transfers_cache")
dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)

season_starts <- 2018:2024      # represent seasons 2018/19 .. 2024/25
sleep_sec <- 3                  # polite pause between requests
max_retries <- 0                # when we ran this for the first time, we had retries. now we set it to 0 to fetch data only from the cache and avoid hitting API calls for missing data
retry_sleep <- 5

out_csv <- file.path(out_dir, "transfers_summer_2018_2025.csv")

# ---------- HELPERS ----------
safe_fn <- function(x) gsub("[^A-Za-z0-9_\\-]", "_", substr(x, 1, 120))

league_urls_cache_file <- function(country_name, start_year) {
  file.path(cache_dir, paste0("league_urls_", safe_fn(country_name), "_", start_year, ".rds"))
}
team_transfers_cache_file <- function(team_url, season_start) {
  file.path(cache_dir, paste0("team_transfers_", safe_fn(team_url), "_", season_start, ".rds"))
}

# cached call for league -> team urls
get_league_team_urls_cached <- function(country_name, start_year) {
  cachef <- league_urls_cache_file(country_name, start_year)
  if (file.exists(cachef)) return(readRDS(cachef))
  for (i in seq_len(max_retries)) {
    try({
      urls <- tm_league_team_urls(country_name = country_name, start_year = start_year)
      saveRDS(urls, cachef)
      Sys.sleep(sleep_sec)
      return(urls)
    }, silent = TRUE)
    message("Retrying league-team-urls for ", country_name, " ", start_year, " (attempt ", i, ")")
    Sys.sleep(retry_sleep)
  }
  warning("Failed to fetch league team URLs for ", country_name, " ", start_year)
  return(NULL)
}

# cached call for team -> transfers (summer)
get_team_transfers_cached <- function(team_url, season_start) {
  cachef <- team_transfers_cache_file(team_url, season_start)
  if (file.exists(cachef)) return(readRDS(cachef))
  for (i in seq_len(max_retries)) {
    try({
      xf <- tm_team_transfers(team_url = team_url, transfer_window = "summer")
      saveRDS(xf, cachef)
      Sys.sleep(sleep_sec)
      return(xf)
    }, silent = TRUE)
    message("Retrying team transfers for ", team_url, " (attempt ", i, ")")
    Sys.sleep(retry_sleep)
  }
  message("No transfers found for team: ", team_url)
  return(NULL)
}

# cheap fuzzy compare to decide whether a returned team name matches a squad in your league table
close_name_match <- function(x, candidates, max_dist = 0.18) {
  # x: single string, candidates: vector of strings
  if (is.na(x) || x == "") return(FALSE)
  # use jaro-winkler via stringdist if available; fallback to grepl substring match
  if (requireNamespace("stringdist", quietly = TRUE)) {
    d <- stringdist::stringdist(tolower(str_squish(x)), tolower(str_squish(candidates)), method = "jw")
    any(d <= max_dist)
  } else {
    any(grepl(tolower(str_squish(x)), tolower(candidates), fixed = TRUE))
  }
}

# ---------- 1) Read league table and derive country+season_start combos ----------
dt <- fread(league_csv)
# use existing 'squad' column (do not create new 'team' column)
if (!"squad" %in% names(dt)) stop("'squad' column not found in league_table.csv")
dt[, squad_clean := tolower(str_squish(squad))]
dt[, season_start := as.integer(sub("([0-9]{4}).*", "\\1", season))]   # extracts starting year like 2018 from "2018/19"
dt <- dt[season_start %in% season_starts]

country_seasons <- unique(dt[, .(country, season_start)])
setorder(country_seasons, country, season_start)

# Build quick lookup of squads per country+season for filtering
squads_by_country_season <- dt[, .(squad_list = list(unique(squad_clean))), by = .(country, season_start)]
setkey(squads_by_country_season, country, season_start)

# ---------- 2) iterate country+season, fetch team URLs, then fetch summer transfers ----------
all_transfers <- list()
for (r in seq_len(nrow(country_seasons))) {
  cs <- country_seasons[r]
  cntry <- cs$country
  sstart <- cs$season_start
  message("Processing: ", cntry, " | season_start ", sstart)
  urls_tbl <- get_league_team_urls_cached(cntry, sstart)
  if (is.null(urls_tbl) || length(urls_tbl) == 0) {
    message(" No team URLs returned for country ", cntry, " season ", sstart, " — skipping")
    next
  }

  # normalize returned structure to a data.table with at least a team_url column and optionally a display name
  if (is.character(urls_tbl)) {
    team_urls_df <- unique(data.table(team_url = urls_tbl))
  } else if (is.data.frame(urls_tbl) || is.data.table(urls_tbl)) {
    # try to detect url column and name column
    url_col <- intersect(names(urls_tbl), c("team_url", "url", "link", "teamlink", "team_url_href"))[1]
    name_col <- intersect(names(urls_tbl), c("team", "club", "club_name", "team_name", "name"))[1]
    if (!is.na(url_col)) {
      team_urls_df <- unique(data.table(team_url = as.character(urls_tbl[[url_col]])))
      if (!is.na(name_col)) team_urls_df[, team_name := as.character(urls_tbl[[name_col]])]
    } else {
      # fallback: use first column as url
      team_urls_df <- unique(data.table(team_url = as.character(urls_tbl[[1]])))
    }
  } else {
    message(" Unrecognized return type from tm_league_team_urls() — skipping this country-season.")
    next
  }

  # OPTIONAL: filter returned URLs to squads that exist in your league_table for that country+season
  # This avoids scraping unrelated clubs. Keep enabled by default.
  filter_local <- TRUE
  if (filter_local && "team_name" %in% names(team_urls_df)) {
    squads_local <- squads_by_country_season[J(cntry, sstart), squad_list][[1]]
    # keep only urls whose team_name roughly matches any squad
    keep_mask <- sapply(team_urls_df$team_name, function(nm) close_name_match(nm, squads_local))
    if (all(!keep_mask)) {
      # if filter removed all, fall back to not filtering (safer)
      message("  Local filtering removed all returned team URLs — falling back to unfiltered list")
      team_urls_final <- unique(team_urls_df$team_url)
    } else {
      team_urls_final <- unique(team_urls_df$team_url[keep_mask])
    }
  } else {
    # if no display name provided by the function, use all returned urls
    team_urls_final <- unique(team_urls_df$team_url)
  }

  # for each team_url, fetch summer transfers (cached)
  for (tu in team_urls_final) {
    message("  Fetching summer transfers for: ", tu)
    xf <- get_team_transfers_cached(tu, sstart)
    if (is.null(xf)) next
    dt_xf <- as.data.table(xf)
    # add context columns so you know which country+season they belong to
    dt_xf[, team_url := tu]
    dt_xf[, country := cntry]
    dt_xf[, season_start := sstart]
    setnames(dt_xf, tolower(names(dt_xf)))
    all_transfers[[length(all_transfers) + 1]] <- dt_xf
  }
}

# ---------- 3) combine and save ----------
if (length(all_transfers) == 0) {
  warning("No transfers fetched. Check that tm_league_team_urls() and tm_team_transfers() work in your environment.")
} else {
  transfers_dt <- rbindlist(all_transfers, use.names = TRUE, fill = TRUE)
  # add season numeric and save
  transfers_dt[, season := season_start]
  fwrite(transfers_dt, out_csv)
  message("Saved summer transfers to: ", out_csv)
  message("Cached per-team rds files are in: ", cache_dir)
}

```

### Filtering Transfers

We add the below filter to ensure that we only have the data for the arrivals in top 5 leagues, and nothing else. (The seller can be from any league. The transfer type need not be permanent.)

```{r}
transfers_path <- "data/transfers_summer_2018_2025.csv"
transfers  <- fread(transfers_path)
league_table <- fread("data/league_table.csv")
top5_leagues <- c("Premier League", "LaLiga", "Bundesliga", "Ligue 1", "Serie A")

top5_league_transfers <- filter(transfers,
       league %in% top5_leagues,
       #league_2 %in% top5_leagues,
       #is_loan == FALSE,
       transfer_type == "Arrivals")
fwrite(top5_league_transfers, "data/top5_league_transfers.csv")
```

### Mapping

Before we proceed with analysis, we try to map the team names between the different tables we have gathered so far.

```{r}

## ===================== PATHS ===================== ##
rivals_path    <- "data/rivalry_continuous.csv"
transfers_path <- "data/top5_league_transfers.csv"
league_path <- "data/league_table.csv"
out_dir     <- "data"

## ===================== 1. LOAD DATA ===================== ##
league <- fread(league_path)
xf     <- fread(transfers_path)

## ===================== 2. BUILD CANONICAL TEAM LIST ===================== ##
# League table defines truth
league[, team_clean := tolower(str_squish(squad))]
league[, country := as.character(country)]

canon_teams <- unique(
  league[, .(country, team_clean)]
)

## ===================== 3. EXTRACT RAW TRANSFER TEAM NAMES ===================== ##
xf[, buyer_raw  := tolower(str_squish(team_name))]
xf[, seller_raw := tolower(str_squish(club_2))]
xf[, country    := as.character(country)]

raw_teams <- unique(rbindlist(list(
  xf[, .(country = country, raw = buyer_raw)],
  xf[, .(country = country_2, raw = seller_raw)]
)))

valid_canon <- canon_teams[, .(country, team_clean)]

raw_teams <- raw_teams[
  country %in% valid_canon$country
]

normalize <- function(x) {
  x |>
    tolower() |>
    str_replace_all("&", "and") |>
    str_replace_all("[^a-z ]", "") |>
    str_squish()
}

valid_canon[, team_norm := normalize(team_clean)]
raw_teams[, raw_norm := normalize(raw)]


## ===================== 4. FUZZY MATCH (RAW → CANONICAL) ===================== ##
team_map <- stringdist_inner_join(
  raw_teams,
  valid_canon,
  by = c("raw" = "team_clean"),
  method = "jw",
  max_dist = 0.2,
  distance_col = "dist"
)

# IMPORTANT: force data.table
setDT(team_map)

## ===================== 5. FIX country.x / country.y ===================== ##
# Keep only same-country matches
team_map <- team_map[country.x == country.y]

# Clean column names
setnames(team_map, "country.x", "country")
team_map[, country.y := NULL]
team_map[, country.dist := NULL]
team_map[, raw.dist := NULL]

# Keep best match per (country, raw)
setorder(team_map, country, raw, dist)
team_map <- team_map[, .SD[1], by = .(country, raw)]
```

We have a very niche problem with selling club's country. unlike a buyer's country that points to the league the country belongs to, seller's country points to the country where the club actually is (Ex: Monaco is not in France, but in the French league). So, when we match based on country, we get NA for seller_clean. After verifying the data, missing values come from such cases. So, we simply use seller_raw's name in seller_clean. The data in seller_raw for these clubs are simplified and don't cause matching issues. So, we can directly copy the name.

```{r}
riv <- fread(rivals_path)   # rivalry_continuous.csv
manual_map <- fread("data/unmatched_transfer_teams.csv")

# Clean rivalry team names
riv[, team_a_clean := tolower(str_squish(team_a))]
riv[, team_b_clean := tolower(str_squish(team_b))]
riv[, season := as.integer(season_start)]

# Keep only needed columns
riv_unique <- unique(
  riv[, .(country, season, team_a_clean, team_b_clean, rivalry_proximity)]
)

# ===================== 2. CLEAN TRANSFER TEAM NAMES ===================== #

team_map_final <- unique(rbind(
  team_map[, .(country, raw, team_clean)],
  manual_map
))

team_map_final[, team_clean := tolower(str_squish(team_clean))]

# Buyer clean
xf <- merge(
  xf,
  team_map_final,
  by.x = c("country", "buyer_raw"),
  by.y = c("country", "raw"),
  all.x = TRUE
)
setnames(xf, "team_clean", "buyer_clean")

xf <- merge(
  xf,
  team_map_final,
  by.x = c("country_2", "seller_raw"),
  by.y = c("country", "raw"),
  all.x = TRUE
)
setnames(xf, "team_clean", "seller_clean")
xf[is.na(seller_clean), seller_clean := seller_raw]

#setnames(xf, "seller_raw", "seller_clean")

# Remove duplicate columns from merges
xf <- xf[, !duplicated(names(xf), fromLast = TRUE), with = FALSE]

# ===================== 3. MAKE TRANSFER PAIRS UNDIRECTED ===================== #

# Create ordered pair to match rivalry file
xf[, team_low  := pmin(buyer_clean, seller_clean)]
xf[, team_high := pmax(buyer_clean, seller_clean)]

# ===================== 4. MERGE CONTINUOUS RIVALRY ===================== #

xf <- merge(
  xf,
  riv_unique,
  by.x = c("country", "season", "team_low", "team_high"),
  by.y = c("country", "season", "team_a_clean", "team_b_clean"),
  all.x = TRUE
)

# ===================== 5. HANDLE CROSS-LEAGUE TRANSFERS ===================== #

# If seller is from a different league (country_2 != country),
# rivalry should be NA explicitly

xf[country_2 != country, rivalry_proximity := NA_real_]

# ===================== 6. CLEANUP ===================== #

xf[, c("team_low", "team_high") := NULL]

xf[, pos_clean := tolower(str_squish(player_position))]
```

We now check for mismatched sellers. We only check for mismatches within the top 5 leagues. We check if the team names in the league table and the seller team's name in transfers match. For other leagues, we don't have league point or rivalry info. So, they don't have to match with the league table.

```{r}
xf
```

```{r}
setDT(league_table)
setDT(xf)

## ---------- NORMALIZATION FUNCTION ----------
normalize <- function(x) {
  x |>
    tolower() |>
    str_replace_all("&", "and") |>
    str_replace_all("[^a-z0-9 ]", "") |>
    str_squish()
}

## ---------- NORMALIZE TEAM NAMES ----------
league_table[, squad_norm := normalize(squad)]
xf[, seller_norm := normalize(seller_clean)]

## ---------- REMOVE DUPLICATES ----------
league_teams <- unique(
  league_table[, .(league_name, squad, squad_norm)]
)

seller_teams <- unique(
  xf[, .(league_name = league_2, seller_clean, seller_norm)]
)

## ---------- FIND MISMATCHES ----------
mismatches <- seller_teams[
  !league_teams,
  on = .(league_name, seller_norm = squad_norm)
]
mismatches
```

In the above data, we can clearly see that none of the mismatched seller teams actually belong to the top 5 leagues. So, we remove this polluted data from transfers.

```{r}
xf <- xf[
  !mismatches,
  on = .(league_2 = league_name, seller_norm)
]
```

Similarly, we check for buyers:

```{r}
xf[, buyer_norm := normalize(buyer_clean)]
buyer_teams <- unique(
  xf[, .(league_name = league, buyer_clean, buyer_norm)]
)

## ---------- FIND MISMATCHES ----------
mismatches <- buyer_teams[
  !league_teams,
  on = .(league_name, buyer_norm = squad_norm)
]
mismatches
```

There are no buyer mismatches.

```{r}
fwrite(xf, "data/top5_league_transfers_clean.csv")
```

## Conditional Multinomial choice model

1.  Choice set: transfers within top 5 leagues (both buyers and sellers are in the top 5 leagues). The buyers of the alternatives are also from the same league as the focal buyer. We also ignore loan transfer options.

```{r}
transfers <- read_csv("data/top5_league_transfers_clean.csv")
rivalry_lookup <- read_csv("data/rivalry_continuous.csv")

transfers_1 <- transfers %>%
  filter(
    league_2 %in% top5_leagues,
    is_loan == FALSE
  ) %>%
  mutate(
    transfer_id = row_number(),
    log_fee = log(transfer_fee + 1)
  )

band_width <- 1

choice_data_1 <- transfers_1 %>%
  rowwise() %>%
  do({
    
    current <- .
    
    focal_log_fee  <- current$log_fee[[1]]
    focal_season   <- current$season[[1]]
    focal_pos      <- current$pos_clean[[1]]
    focal_country  <- current$country[[1]]
    focal_league  <- current$league[[1]]
    focal_id       <- current$transfer_id[[1]]
    
    pool <- transfers_1 %>%
      filter(
        season == focal_season,
        pos_clean == focal_pos,
        league == focal_league,
        transfer_id != focal_id,
        !is.na(log_fee),
        abs(log_fee - focal_log_fee) <= band_width
      )
    
    pool <- bind_rows(pool, current)
    
    pool %>%
      mutate(
        choice_id = focal_id,
        nat_match = ifelse(player_nationality == focal_country, 1, 0),
        chosen = ifelse(transfer_id == focal_id, 1, 0)
      )
    
  }) %>%
  ungroup()

choice_data_1 <- choice_data_1 %>%
  mutate(
    rivalry_proximity = ifelse(is.na(rivalry_proximity), 0, rivalry_proximity)
  )
```

```{r}
#remove choice sets where there are no alternatives
setDT(choice_data_1)
choice_data_1 <- choice_data_1[
  choice_id %in% choice_data_1[, .N, by = choice_id][N > 1, choice_id]
]

fwrite(choice_data_1,"data/choice_data.csv")
choice_data_1
```

```{r}

model_1 <- clogit(
  chosen ~ player_age +
    minutes_played + 
    nat_match +
    goals +
    log_fee + 
    rivalry_proximity + 
    strata(choice_id),
  data = choice_data_1
)

summary(model_1)
```

2.  All arrivals to top 5 leagues (including loans). The Seller can be from any league. We also choose the top 10 alternatives for each transfer.

```{r}
transfers <- transfers %>%
  mutate(
    transfer_id = row_number(),
    log_fee = log(transfer_fee + 1)
  )

set.seed(123)
band_width <- 1
K <- 10

choice_data_2 <- transfers %>%
  rowwise() %>%
  do({
    
    current <- .
    
    focal_log_fee  <- current$log_fee[[1]]
    focal_season   <- current$season[[1]]
    focal_country  <- current$country[[1]]
    focal_league  <- current$league[[1]]
    focal_pos      <- current$pos_clean[[1]]
    focal_id       <- current$transfer_id[[1]]
    
    # Candidate alternatives
    pool_all <- transfers %>%
      filter(
        season == focal_season,
        league == focal_league,
        pos_clean == focal_pos,
        transfer_id != focal_id,
        !is.na(log_fee),
        abs(log_fee - focal_log_fee) <= band_width
      )
    
    # Randomly sample up to K alternatives
    if (nrow(pool_all) > K) {
      pool_sample <- pool_all %>% slice_sample(n = K)
    } else {
      pool_sample <- pool_all
    }
    
    # Add chosen transfer back
    pool <- bind_rows(pool_sample, current)
    
    pool %>%
      mutate(
        choice_id = focal_id,
        nat_match = ifelse(player_nationality == focal_country, 1, 0),
        chosen = ifelse(transfer_id == focal_id, 1, 0)
      )
    
  }) %>%
  ungroup()

choice_data_2 <- choice_data_2 %>%
  mutate(
    rivalry_proximity = ifelse(is.na(rivalry_proximity), 0, rivalry_proximity)
  )

setDT(choice_data_2)
choice_data_2 <- choice_data_2[
  choice_id %in% choice_data_2[, .N, by = choice_id][N > 1, choice_id]
]
```

```{r}
model_2 <- clogit(
  chosen ~ player_age +
    minutes_played + nat_match +
    goals +
    log_fee + 
    rivalry_proximity + 
    strata(choice_id),
  data = choice_data_2
)

summary(model_2)
```

3.  Either the alternative buyer is in the focal league, or the alternative seller is

```{r}
transfers <- read_csv("data/top5_league_transfers_clean.csv")
transfers <- transfers %>%
  filter(
    is_loan == FALSE
  ) %>%
  mutate(
    transfer_id = row_number(),
    log_fee = log(transfer_fee + 1)
  )

set.seed(123)
band_width <- 1
K <- 10

choice_data_3 <- transfers %>%
  rowwise() %>%
  do({
    
    current <- .
    
    focal_log_fee  <- current$log_fee[[1]]
    focal_season   <- current$season[[1]]
    focal_country  <- current$country[[1]]
    focal_league  <- current$league[[1]]
    focal_pos      <- current$pos_clean[[1]]
    focal_id       <- current$transfer_id[[1]]
    
    # Candidate alternatives
    pool_all <- transfers %>%
      filter(
        season == focal_season,
        league == focal_league | league_2 == focal_league,
        pos_clean == focal_pos,
        transfer_id != focal_id,
        !is.na(log_fee),
        abs(log_fee - focal_log_fee) <= band_width
      )
    
    # Randomly sample up to K alternatives
    if (nrow(pool_all) > K) {
      pool_sample <- pool_all %>% slice_sample(n = K)
    } else {
      pool_sample <- pool_all
    }
    
    # Add chosen transfer back
    pool <- bind_rows(pool_sample, current)
    
    pool %>%
      mutate(
        choice_id = focal_id,
        nat_match = ifelse(player_nationality == focal_country, 1, 0),
        chosen = ifelse(transfer_id == focal_id, 1, 0)
      )
    
  }) %>%
  ungroup()

choice_data_3 <- choice_data_3 %>%
  mutate(
    rivalry_proximity = ifelse(is.na(rivalry_proximity), 0, rivalry_proximity)
  )

setDT(choice_data_3)
choice_data_3 <- choice_data_3[
  choice_id %in% choice_data_3[, .N, by = choice_id][N > 1, choice_id]
]

fwrite(choice_data_3, "data/choice_data_3.csv")
```

```{r}
model_3 <- clogit(
  chosen ~ player_age +
    minutes_played + nat_match +
    goals +
    log_fee + 
    rivalry_proximity + 
    strata(choice_id),
  data = choice_data_3
)

summary(model_3)
```

A 1-unit increase in rivalry proximity multiplies the odds of choosing that player by \~2.47. So clubs are much more likely to sign players from rivals. The effect is also very strong.

## Analysis

The first 2 models produce very high p-values for every variable. Also, the coefficients are close to 0 and the concordance value of 0.5 and LR, Wald and Score tests have value ≈ 0. That combination means the model cannot distinguish chosen from non-chosen alternatives. So, we analyse the data to check if there are any issues with the choice set.

We already know that team names match. So there are no issues there. We first check how different the alternative choices are:

```{r}
setDT(choice_data_1)
sd_table <- choice_data_1[, .(
  age_sd = sd(player_age),
  minutes_sd = sd(minutes_played),
  goals_sd = sd(goals),
  fee_sd = sd(log_fee),
  riv_sd = sd(rivalry_proximity)
), by = choice_id]

sd_counts <- sd_table[, .(
  age_sd_zero     = sum(age_sd == 0, na.rm = TRUE),
  age_sd_nonzero  = sum(age_sd > 0, na.rm = TRUE),

  minutes_sd_zero    = sum(minutes_sd == 0, na.rm = TRUE),
  minutes_sd_nonzero = sum(minutes_sd > 0, na.rm = TRUE),

  goals_sd_zero    = sum(goals_sd == 0, na.rm = TRUE),
  goals_sd_nonzero = sum(goals_sd > 0, na.rm = TRUE),

  fee_sd_zero    = sum(fee_sd == 0, na.rm = TRUE),
  fee_sd_nonzero = sum(fee_sd > 0, na.rm = TRUE),

  riv_sd_zero    = sum(riv_sd == 0, na.rm = TRUE),
  riv_sd_nonzero = sum(riv_sd > 0, na.rm = TRUE)
)]

sd_counts
```

As we can see, there are a few instances when the alternatives don't have much difference between them, but there is still plenty of data where the alternatives aren't too similar. So, this can't be the issue.

We now check how often the chosen alternative is the “best” option within its choice set according to each variable, and average across choice sets. If the values ≈ random chance, the variable does not explain the choice.

```{r}
choice_data_1[, .(

  age_best = if (all(is.na(player_age))) NA_real_
  else mean(chosen == 1 & player_age == min(player_age, na.rm = TRUE)),

  minutes_best = if (all(is.na(minutes_played))) NA_real_
  else mean(chosen == 1 & minutes_played == max(minutes_played, na.rm = TRUE)),

  goals_best = if (all(is.na(goals))) NA_real_
  else mean(chosen == 1 & goals == max(goals, na.rm = TRUE)),

  fee_best = if (all(is.na(log_fee))) NA_real_
  else mean(chosen == 1 & log_fee == min(log_fee, na.rm = TRUE)),

  riv_best = if (all(is.na(rivalry_proximity))) NA_real_
  else mean(chosen == 1 & rivalry_proximity == max(rivalry_proximity, na.rm = TRUE))

), by = choice_id][, lapply(.SD, function(x) mean(x, na.rm = TRUE))]
```

```{r}
choice_data_1[, mean(.N), by = choice_id][, mean(V1)]
```

We can see that the average choice set size is about 3. By random selection, the best player would be selected 33% of the time. But from our data, the predictors are actually worse than random at predicting the chosen player, which explains why the Cox model estimates coefficients ≈ 0.

Next, we check the average rank of the chosen player within each choice set for each variable, averaged across choice sets. If the average rank is better than random, the variable helps explain the choice; if it equals the random rank, the variable has no predictive power.

```{r}
choice_data_1[, .(

  age_rank     = if (all(is.na(player_age))) NA_real_
                 else mean(frank(player_age, ties.method="average") [chosen==1]),

  minutes_rank = if (all(is.na(minutes_played))) NA_real_
                 else mean(frank(-minutes_played, ties.method="average") [chosen==1]),

  goals_rank   = if (all(is.na(goals))) NA_real_
                 else mean(frank(-goals, ties.method="average") [chosen==1]),

  fee_rank     = if (all(is.na(log_fee))) NA_real_
                 else mean(frank(log_fee, ties.method="average") [chosen==1]),

  riv_rank     = if (all(is.na(rivalry_proximity))) NA_real_
                 else mean(frank(-rivalry_proximity, ties.method="average") [chosen==1])

), by = choice_id][, lapply(.SD, mean, na.rm = TRUE)]
```

These are almost exactly equal to the random benchmark (2.215). Thus, the chosen player is indistinguishable from random. Thus, the data simply contains no predictive signal for the choice in the variables that were included.

Unlike the first 2 models, the 3rd one actually produced a very low p-value for rivalry_proximity, and the LR, Wald and Score test showed about 58%. So, let's try the same analysis on choice_data_3:

```{r}
choice_data_3[, .(

  age_best = if (all(is.na(player_age))) NA_real_
  else mean(chosen == 1 & player_age == min(player_age, na.rm = TRUE)),

  minutes_best = if (all(is.na(minutes_played))) NA_real_
  else mean(chosen == 1 & minutes_played == max(minutes_played, na.rm = TRUE)),

  goals_best = if (all(is.na(goals))) NA_real_
  else mean(chosen == 1 & goals == max(goals, na.rm = TRUE)),

  fee_best = if (all(is.na(log_fee))) NA_real_
  else mean(chosen == 1 & log_fee == min(log_fee, na.rm = TRUE)),

  riv_best = if (all(is.na(rivalry_proximity))) NA_real_
  else mean(chosen == 1 & rivalry_proximity == max(rivalry_proximity, na.rm = TRUE))

), by = choice_id][, lapply(.SD, function(x) mean(x, na.rm = TRUE))]
```

```{r}
choice_data_3[, mean(.N), by = choice_id][, mean(V1)]
```

The results are similar to choice_data_1

```{r}
choice_data_3[, .(

  age_rank     = if (all(is.na(player_age))) NA_real_
                 else mean(frank(player_age, ties.method="average") [chosen==1]),

  minutes_rank = if (all(is.na(minutes_played))) NA_real_
                 else mean(frank(-minutes_played, ties.method="average") [chosen==1]),

  goals_rank   = if (all(is.na(goals))) NA_real_
                 else mean(frank(-goals, ties.method="average") [chosen==1]),

  fee_rank     = if (all(is.na(log_fee))) NA_real_
                 else mean(frank(log_fee, ties.method="average") [chosen==1]),

  riv_rank     = if (all(is.na(rivalry_proximity))) NA_real_
                 else mean(frank(-rivalry_proximity, ties.method="average") [chosen==1])

), by = choice_id][, lapply(.SD, mean, na.rm = TRUE)]
```

Again, the results are similar.
