---
title: "What motivates a transfer: player performance, transfer fee or the rivalry between the teams?"
author: 
  - name: "Vignesh Thiruvel Selvam"
  - name: "Prof. Dr.Christoph Ihl"
    role: "Supervisor"
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
execute:
  echo: false
editor: visual
---

# Introduction

Teams in the Top 5 Men's Leagues of Europe (Premier League, La Liga, Bundesliga, Serie A, Ligue 1) are known for their big spending every Summer transfer window. A team that is newly crowned Champion would buy players to continue on their fine form. Teams that got newly promoted would look at options to extend their stay in the top flight. Other teams would try to find players to improve their performance and build upon last season's results. Thus, no matter where a team finishes, once the Summer transfer window opens, all teams are active in the market. But how do they choose a player? Is it purely based on numbers (stats, age, fee)? Or could there be bigger forces at play, like a player's nationality or even the rivalry between clubs in a league?

In order to answer this question, we compile a list of teams in the Top 5 leagues in Europe, and their transfer history (only players bought) from the Summer transfer window in 2018 to 2024. We then combine it with a team's ELO and define a rivalry score. We also make use of a player's attributes such as their age, nationality, minutes played, etc. along with their transfer fee to determine the motivating factor of a transfer.

```{r message = FALSE, warning = FALSE}
library(tidyverse)
library(worldfootballR)
library(dplyr)
library(data.table)
library(stringr)
library(readr)
library(fuzzyjoin)
library(survival)
library(modelsummary)
library(broom)
library(knitr)
library(kableExtra)
```

# Data Collection

We obtain league tables and ELO ratings for each season from the following sources:

-   ELO: [ClubELO](http://clubelo.com/API) API

-   League tables and Transfers: Scraped from [Transfermarkt](https://www.transfermarkt.com/) using [WorldFootballR](https://jaseziv.github.io/worldfootballR/) package

The league table for each league gives us the final standings of the teams at the end of each season in our observed range. It gives us the position the team finished in and the points the team has scored that season.

From the ClubELO API, we can get a list of ELO ratings of teams for a given date. We gather the ELO data for the end of each season. We use the team names from both these tables and create a combined table (with points and ELO as required for [rivalry measurement]) later.

```{r include = FALSE}

# ------1. Getting League Data------
league_path <- "data/league_table.csv"
if (!file.exists(league_path)) {
  leagues <- c("England", "Spain", "Germany", "Italy", "France")
  seasons <- 2017:2024
  league_season_combos <- expand.grid(league = leagues, season = seasons)
  
  # Parallel fetching
  results <- future_pmap_dfr(league_season_combos, function(league, season) {
    tryCatch({
      matchday <- case_when(
        league == "Germany" ~ 34,
        league == "France" & season == 2019 ~ 27,
        league == "France" & season >= 2023 ~ 34,
        TRUE ~ 38
      )
      message(paste("Fetching:", league, season))
      df <- tm_matchday_table(country_name=league, start_year=season, matchday=matchday)
      df$season <- season
      df
    }, error = function(e) {
      message(paste("Error fetching", league, season, ":", e$message))
      return(NULL)
    })
  })
  results[, league_name := league |> str_remove("Games of Matchday \\d+ ") |> str_remove(" \\d{2}/\\d{2}")]
  fwrite(results, league_path)
}
```

```{r include = FALSE}
# 1) read
dt_raw <- fread("data/league_table.csv")

dt_raw[, team := str_squish(squad)]
dt_raw[, season := as.character(season)]
dt_raw[, league := as.character(league)]
dt_raw[, pos := as.integer(rk)]

if ("matchday" %in% names(dt_raw)) {
  dt_final <- dt_raw[!is.na(pos), .SD[which.max(as.numeric(matchday))], 
                     by = .(country, league, season, team)]
} else {
  dt_final <- dt_raw[!is.na(pos), .SD[1], by = .(country, league, season, team)]
}
# Keep only needed cols
dt_final <- dt_final[, .(country, league, season, team, pos, pts)]
dt_final[, team_low := tolower(team)]
dt_final[, country_low := tolower(country)]
dt_final[, league_low := tolower(league)]
```

```{r include = FALSE}
# ------2. Getting ELO------

# ---------------- PARAMETERS ----------------
seasons <- 2018:2025
snapshot_month_day <- "-09-01"
max_name_dist <- 0.18

# ---------------- PATHS (uses uploaded files in data) ----------------
league_csv <- "data/league_table.csv"
manual_map_path <- "data/manual_club_map.csv"
out_dir <- "data"
out_rivals_csv <- file.path(out_dir, "rivalry_lookup.csv")
out_elo_raw_dir <- file.path(out_dir, "elo_snapshots")
dir.create(out_elo_raw_dir, showWarnings = FALSE, recursive = TRUE)

# ---------------- 1) load & prepare league table (dt_final) ----------------
dt_raw <- fread(league_csv)
dt_raw[, team := str_squish(squad)]
dt_raw[, country := as.character(country)]
dt_raw[, league := as.character(league)]
dt_raw[, season := as.character(season)]
dt_raw[, pos := as.integer(rk)]

if ("matchday" %in% names(dt_raw)) {
  dt_final <- dt_raw[!is.na(pos), .SD[which.max(as.numeric(matchday))], by = .(country, league, season, team)]
} else {
  dt_final <- dt_raw[!is.na(pos), .SD[1], by = .(country, league, season, team)]
}
dt_final <- dt_final[, .(country, league, season, team, pos)]
dt_final[, team_clean := tolower(str_squish(team))]
# Precompute season_start string on dt_final for lookups (e.g., "2017" from "2017/18")
dt_final[, season_start_str := sub("([0-9]{4}).*", "\\1", season)]

# ---------------- 2) fetch ClubElo snapshots (cached) ----------------
fetch_elo_for_date <- function(date_str) {
  destfile <- file.path(out_elo_raw_dir, paste0("clubelo_", date_str, ".csv"))
  if (file.exists(destfile)) {
    message("Using cached Elo snapshot: ", destfile)
    df <- read_csv(destfile, show_col_types = FALSE)
    setDT(df); return(df)
  }
  url <- paste0("http://api.clubelo.com/", date_str)
  message("Fetching from API: ", url)
  r <- httr::GET(url)
  if (httr::status_code(r) != 200) stop("ClubElo request failed for date ", date_str, " (status ", httr::status_code(r), ")")
  txt <- httr::content(r, as = "text", encoding = "UTF-8")
  df <- read_csv(txt, show_col_types = FALSE)
  setDT(df)
  setnames(df, tolower(names(df)))
  club_col <- grep("club|team|name", names(df), value = TRUE)[1]
  elo_col  <- grep("elo|rating", names(df), value = TRUE)[1]
  if (is.na(club_col) || is.na(elo_col)) stop("Unable to find club/elo columns in ClubElo CSV for date: ", date_str)
  df <- df[, .(club = get(club_col), elo = as.numeric(get(elo_col)))]
  df <- df[!is.na(elo) & !is.na(club)]
  df[, club_clean := tolower(str_squish(gsub("\\(.*\\)", "", club)))]
  df <- df[order(-elo)]
  df[, snapshot_date := date_str]
  write_csv(df, destfile)
  message("Saved snapshot: ", destfile)
  return(df)
}

# Fetch and cache all snapshots (only does network calls if missing)
elo_all <- rbindlist(lapply(seasons, function(y) {
  d <- paste0(y, snapshot_month_day)
  message("Fetch snapshot: ", d)
  Sys.sleep(0.5) # be kind to API
  fetch_elo_for_date(d)
}), use.names = TRUE, fill = TRUE)
message("Total Elo rows fetched (or cached): ", nrow(elo_all))
elo_all[, season_start := as.character(as.integer(substr(snapshot_date, 1, 4)) - 1L)]

# ---------------- 3) APPLY MANUAL MAPPINGS (rowwise, robust) ----------------
manual_map_applied <- data.table()  # result holder
if (file.exists(manual_map_path)) {
  mm <- fread(manual_map_path)
  # normalize columns: support snapshot_date or season_start
  if (!("snapshot_date" %in% names(mm)) & ("season_start" %in% names(mm))) {
    mm[, snapshot_date := paste0(season_start, snapshot_month_day)]
  }
  if (!("club" %in% names(mm)) | !("mapped_team" %in% names(mm))) {
    stop("manual_club_map.csv must contain at minimum: club, mapped_team, and snapshot_date or season_start")
  }
  mm[, club_clean := tolower(str_squish(gsub("\\(.*\\)", "", club)))]
  mm[, mapped_team_clean := tolower(str_squish(mapped_team))]
  mm[, season_start := as.character(substring(snapshot_date, 1, 4))]
  # rowwise validate and build manual_map_applied
  valid_rows <- list()
  for (i in seq_len(nrow(mm))) {
    row <- mm[i]
    ss <- as.character(row$season_start)
    mt_clean <- row$mapped_team_clean
    # check dt_final: does mapped team exist in that season_start?
    cand <- dt_final[team_clean == mt_clean & season_start_str == ss]
    if (nrow(cand) == 0L) next  # invalid mapping: mapped_team not present in league for that season
    # find matching elo rows for the same snapshot + club_clean
    erows <- elo_all[season_start == ss & club_clean == row$club_clean]
    if (nrow(erows) == 0L) next  # no matching elo row to attach
    # build manual match rows
    for (er in seq_len(nrow(erows))) {
      for (cd in seq_len(nrow(cand))) {
        valid_rows[[length(valid_rows) + 1]] <- data.table(
          snapshot_date = erows[er, snapshot_date],
          club = erows[er, club],
          club_clean = erows[er, club_clean],
          elo = erows[er, elo],
          team = cand[cd, team],
          team_clean = cand[cd, team_clean],
          country = cand[cd, country],
          league = cand[cd, league],
          dist = 0  # mark manual mapping as exact
        )
      }
    }
  }
  if (length(valid_rows) > 0) {
    manual_map_applied <- rbindlist(valid_rows, use.names = TRUE, fill = TRUE)
    message("Applied ", nrow(manual_map_applied), " manual mapping rows from: ", manual_map_path)
  } else {
    message("No valid manual mappings found in: ", manual_map_path)
  }
} else {
  message("No manual_club_map.csv found at: ", manual_map_path, " (continuing without manual mappings)")
}

# ---------------- 4) FUZZY MATCH remaining Elo clubs -> dt_final teams ----------------
# Build candidate list excluding manual-mapped (by snapshot_date+club_clean)
elo_to_match <- copy(elo_all[, .(snapshot_date, club, club_clean, elo, season_start)])
if (nrow(manual_map_applied) > 0) {
  keys_mm <- unique(manual_map_applied[, .(snapshot_date, club_clean)])
  setkey(elo_to_match, snapshot_date, club_clean)
  setkey(keys_mm, snapshot_date, club_clean)

  # Proper anti-join (no need for fsetdiff)
  elo_to_match <- elo_to_match[!keys_mm]
}

# Build teams by season_start
teams_by_start <- unique(dt_final[, .(season_start_str, team, team_clean, country, league)])
setnames(teams_by_start, "season_start_str", "season_start")

matched_list <- list()
unmatched_list <- list()
for (yr in sort(unique(elo_to_match$season_start))) {
  eloy <- elo_to_match[season_start == yr, .(snapshot_date, club, club_clean, elo)]
  targ <- teams_by_start[season_start == yr, .(team, team_clean, country, league)]
  if (nrow(targ) == 0L) { unmatched_list[[yr]] <- eloy; next }
  mj <- stringdist_inner_join(eloy, targ, by = c("club_clean" = "team_clean"),
                              method = "jw", max_dist = max_name_dist, distance_col = "dist")
  setDT(mj); setorder(mj, snapshot_date, club, dist)
  if (nrow(mj) > 0) {
    mj <- mj[, .SD[1], by = .(snapshot_date, club)]  # best match per elo club
    matched_list[[yr]] <- mj
    matched_clubs <- unique(mj$club)
    unmatched_list[[yr]] <- eloy[!club %in% matched_clubs]
  } else {
    unmatched_list[[yr]] <- eloy
  }
}

matched_dt_fuzzy <- if (length(matched_list) > 0) rbindlist(matched_list, use.names = TRUE, fill = TRUE) else data.table()
unmatched_dt <- if (length(unmatched_list) > 0) rbindlist(unmatched_list, use.names = TRUE, fill = TRUE) else data.table()

# Combine manual-applied matches and fuzzy matches into matched_dt
if (nrow(manual_map_applied) > 0 && nrow(matched_dt_fuzzy) > 0) {
  matched_dt <- rbindlist(list(manual_map_applied, matched_dt_fuzzy[, .(snapshot_date, club, club_clean, elo, team, team_clean, country, league, dist)]),
                          use.names = TRUE, fill = TRUE)
} else if (nrow(manual_map_applied) > 0) {
  matched_dt <- copy(manual_map_applied)
} else {
  matched_dt <- copy(matched_dt_fuzzy)
}

# Save remaining unmatched for inspection
if (nrow(unmatched_dt) > 0) {
  fwrite(unique(unmatched_dt), file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
  message("Wrote remaining unmatched Elo clubs to: ", file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
} else {
  # produce empty file to indicate nothing remaining
  fwrite(data.table(snapshot_date=character(), club=character(), club_clean=character(), elo=numeric()),
         file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
  message("No remaining unmatched Elo clubs; created empty file: ", file.path(out_dir, "elo_unmatched_clubs_remaining.csv"))
}

# ---------------- 5) set Elo-season = snapshot_year - 1 (numeric) for Elo outputs ONLY ----------------
if (nrow(matched_dt) == 0) {
  message("Warning: matched_dt is empty (no Elo->league matches). Elo-only will be empty.")
}
matched_dt[, snapshot_year := as.integer(substr(snapshot_date, 1, 4))]
matched_dt[, elo_season := snapshot_year - 1L]  # e.g., 2018 -> 2017
```

The transfer data consists of players bought by all teams in the top 5 leagues from 2018 to 2024 summer transfer windows. This includes both permanent moves and loan moves (in football, at the end of a loan deal, a player returns to their parent club).

In football, there are 2 transfer windows each season when the teams are allowed to buy players. The first one is the Summer transfer window, usually between June to the end of August before the start of a season. The second window, commonly known as the winter transfer window is usually in the middle of a season in January. For this research we don't consider transfers in the winter window because:

1.  Winter transfers are based on a team's position in the middle of a season. Often, it is considered to be a temporary solution (to fill up an injured player's spot or a desperate transfer to avoid relegation, etc.).
2.  It could even affect a team's options in the next summer window, due to constrained budget (already spent on players in the winter transfer), overcrowded positions (a player was bought due to an injury to an existing player, but the injured player has now returned, etc.).

Thus, it is a completely different strategy compared to summer transfers and beyond the scope of what we are trying to establish in this analysis.

In addition to the selling club, buying club, their leagues and transfer fee, we also get information about the player such as their age, number of goals scored, minutes played last season, etc. that we use for our analysis.

```{r include = FALSE}
# ------3. Getting Transfers------

# ---------- PARAMETERS & PATHS ----------
league_csv <- "data/league_table.csv"   # your uploaded file (use existing 'squad' col)
out_dir <- "data"
cache_dir <- file.path(out_dir, "transfers_cache")
dir.create(cache_dir, recursive = TRUE, showWarnings = FALSE)

season_starts <- 2018:2024      # represent seasons 2018/19 .. 2024/25
sleep_sec <- 3                  # polite pause between requests
max_retries <- 0                # when we executed this for the first time, we had retries. now we set it to 0 to fetch data only from the cache and avoid hitting API calls for missing data
retry_sleep <- 5

out_csv <- file.path(out_dir, "transfers_summer_2018_2025.csv")

# ---------- HELPERS ----------
safe_fn <- function(x) gsub("[^A-Za-z0-9_\\-]", "_", substr(x, 1, 120))

league_urls_cache_file <- function(country_name, start_year) {
  file.path(cache_dir, paste0("league_urls_", safe_fn(country_name), "_", start_year, ".rds"))
}
team_transfers_cache_file <- function(team_url, season_start) {
  file.path(cache_dir, paste0("team_transfers_", safe_fn(team_url), "_", season_start, ".rds"))
}

# cached call for league -> team urls
get_league_team_urls_cached <- function(country_name, start_year) {
  cachef <- league_urls_cache_file(country_name, start_year)
  if (file.exists(cachef)) return(readRDS(cachef))
  for (i in seq_len(max_retries)) {
    try({
      urls <- tm_league_team_urls(country_name = country_name, start_year = start_year)
      saveRDS(urls, cachef)
      Sys.sleep(sleep_sec)
      return(urls)
    }, silent = TRUE)
    message("Retrying league-team-urls for ", country_name, " ", start_year, " (attempt ", i, ")")
    Sys.sleep(retry_sleep)
  }
  warning("Failed to fetch league team URLs for ", country_name, " ", start_year)
  return(NULL)
}

# cached call for team -> transfers (summer)
get_team_transfers_cached <- function(team_url, season_start) {
  cachef <- team_transfers_cache_file(team_url, season_start)
  if (file.exists(cachef)) return(readRDS(cachef))
  for (i in seq_len(max_retries)) {
    try({
      xf <- tm_team_transfers(team_url = team_url, transfer_window = "summer")
      saveRDS(xf, cachef)
      Sys.sleep(sleep_sec)
      return(xf)
    }, silent = TRUE)
    message("Retrying team transfers for ", team_url, " (attempt ", i, ")")
    Sys.sleep(retry_sleep)
  }
  message("No transfers found for team: ", team_url)
  return(NULL)
}

# cheap fuzzy compare to decide whether a returned team name matches a squad in your league table
close_name_match <- function(x, candidates, max_dist = 0.18) {
  # x: single string, candidates: vector of strings
  if (is.na(x) || x == "") return(FALSE)
  # use jaro-winkler via stringdist if available; fallback to grepl substring match
  if (requireNamespace("stringdist", quietly = TRUE)) {
    d <- stringdist::stringdist(tolower(str_squish(x)), tolower(str_squish(candidates)), method = "jw")
    any(d <= max_dist)
  } else {
    any(grepl(tolower(str_squish(x)), tolower(candidates), fixed = TRUE))
  }
}

# ---------- 1) Read league table and derive country+season_start combos ----------
dt <- fread(league_csv)
# use existing 'squad' column (do not create new 'team' column)
if (!"squad" %in% names(dt)) stop("'squad' column not found in league_table.csv")
dt[, squad_clean := tolower(str_squish(squad))]
dt[, season_start := as.integer(sub("([0-9]{4}).*", "\\1", season))]   # extracts starting year like 2018 from "2018/19"
dt <- dt[season_start %in% season_starts]

country_seasons <- unique(dt[, .(country, season_start)])
setorder(country_seasons, country, season_start)

# Build quick lookup of squads per country+season for filtering
squads_by_country_season <- dt[, .(squad_list = list(unique(squad_clean))), by = .(country, season_start)]
setkey(squads_by_country_season, country, season_start)

# ---------- 2) iterate country+season, fetch team URLs, then fetch summer transfers ----------
all_transfers <- list()
for (r in seq_len(nrow(country_seasons))) {
  cs <- country_seasons[r]
  cntry <- cs$country
  sstart <- cs$season_start
  message("Processing: ", cntry, " | season_start ", sstart)
  urls_tbl <- get_league_team_urls_cached(cntry, sstart)
  if (is.null(urls_tbl) || length(urls_tbl) == 0) {
    message(" No team URLs returned for country ", cntry, " season ", sstart, " — skipping")
    next
  }

  # normalize returned structure to a data.table with at least a team_url column and optionally a display name
  if (is.character(urls_tbl)) {
    team_urls_df <- unique(data.table(team_url = urls_tbl))
  } else if (is.data.frame(urls_tbl) || is.data.table(urls_tbl)) {
    # try to detect url column and name column
    url_col <- intersect(names(urls_tbl), c("team_url", "url", "link", "teamlink", "team_url_href"))[1]
    name_col <- intersect(names(urls_tbl), c("team", "club", "club_name", "team_name", "name"))[1]
    if (!is.na(url_col)) {
      team_urls_df <- unique(data.table(team_url = as.character(urls_tbl[[url_col]])))
      if (!is.na(name_col)) team_urls_df[, team_name := as.character(urls_tbl[[name_col]])]
    } else {
      # fallback: use first column as url
      team_urls_df <- unique(data.table(team_url = as.character(urls_tbl[[1]])))
    }
  } else {
    message(" Unrecognized return type from tm_league_team_urls() — skipping this country-season.")
    next
  }

  # OPTIONAL: filter returned URLs to squads that exist in your league_table for that country+season
  # This avoids scraping unrelated clubs. Keep enabled by default.
  filter_local <- TRUE
  if (filter_local && "team_name" %in% names(team_urls_df)) {
    squads_local <- squads_by_country_season[J(cntry, sstart), squad_list][[1]]
    # keep only urls whose team_name roughly matches any squad
    keep_mask <- sapply(team_urls_df$team_name, function(nm) close_name_match(nm, squads_local))
    if (all(!keep_mask)) {
      # if filter removed all, fall back to not filtering (safer)
      message("  Local filtering removed all returned team URLs — falling back to unfiltered list")
      team_urls_final <- unique(team_urls_df$team_url)
    } else {
      team_urls_final <- unique(team_urls_df$team_url[keep_mask])
    }
  } else {
    # if no display name provided by the function, use all returned urls
    team_urls_final <- unique(team_urls_df$team_url)
  }

  # for each team_url, fetch summer transfers (cached)
  for (tu in team_urls_final) {
    message("  Fetching summer transfers for: ", tu)
    xf <- get_team_transfers_cached(tu, sstart)
    if (is.null(xf)) next
    dt_xf <- as.data.table(xf)
    # add context columns so you know which country+season they belong to
    dt_xf[, team_url := tu]
    dt_xf[, country := cntry]
    dt_xf[, season_start := sstart]
    setnames(dt_xf, tolower(names(dt_xf)))
    all_transfers[[length(all_transfers) + 1]] <- dt_xf
  }
}

# ---------- 3) combine and save ----------
if (length(all_transfers) == 0) {
  warning("No transfers fetched. Check that tm_league_team_urls() and tm_team_transfers() work in your environment.")
} else {
  transfers_dt <- rbindlist(all_transfers, use.names = TRUE, fill = TRUE)
  # add season numeric and save
  transfers_dt[, season := season_start]
  fwrite(transfers_dt, out_csv)
  message("Saved summer transfers to: ", out_csv)
  message("Cached per-team rds files are in: ", cache_dir)
}
```

We then need to filter the transfers to ensure that we only have the data for the arrivals in top 5 leagues (The seller can be from any league. The transfer type need not be permanent). Redundant data also includes the return of a player to his parent club at the end of a loan spell. We remove this as well.

```{r include = FALSE}
# ------4. Filtering transfers------
transfers_path <- "data/transfers_summer_2018_2025.csv"
transfers  <- fread(transfers_path)
league_table <- fread("data/league_table.csv")
top5_leagues <- c("Premier League", "LaLiga", "Bundesliga", "Ligue 1", "Serie A")

top5_league_transfers <- filter(transfers,
       league %in% top5_leagues,
       transfer_type == "Arrivals")

top5_league_transfers <- top5_league_transfers %>%
  mutate(
    transfer_fee = case_when(
      is_loan & str_detect(transfer_notes, "^End of loan") ~ NA_real_,
      is_loan ~ 0,
      TRUE ~ transfer_fee
    )
  )
fwrite(top5_league_transfers, "data/top5_league_transfers.csv")
```

# Rivalry Measurement

We define competitive rivalry based on the following strategy:

1.  We only measure rivalry between teams that belong to the same league. Teams from different leagues do compete with each other in certain competitions, but are not necessarily considered rivals. Thus, rivalry is meaningful only within a league.

2.  Next, we define a vector:

    $$
    \mathbf{x}_{i,t} =\left(z^{(t)}_{i,t},\;z^{(t-1)}_{i,t},\;z^{(t-2)}_{i,t},\;z^{(\mathrm{elo})}_{i,t}\right)
    $$

    where:

    $x_{i,t}$ is the feature vector for team $i$ in season $t$, consisting of standardized (z-scored) performance measures\
    $z^{(t)}_{i,t}$: is the points in season $t$\
    $z^{(t-1)}_{i,t}$: is the points in season $t-1$\
    $z^{(t-2)}_{i,t}$: is the points in season $t-2$\
    $z^{(\mathrm{elo})}_{i,t}$: is the ELO rating at the start of season $t$\
    $d_{ij,t}$: is the rivalry distance between teams $i$ and $j$ in season $t$\

3.  Using this vector, we calculate distances between the teams based on a Modified Euclidean approach.

    $$
    d_{ij,t} =\sqrt{\frac{1}{K_{ij,t}}\sum_{k \in \mathcal{K}_{ij,t}}\left(x_{i,k,t} - x_{j,k,t}\right)^2}
    $$

    where:

    $d_{ij,t}$ is the distance between teams $i$ and $j$ in season $t$\
    $\mathcal{K}_{ij,t}$ is the set of valid dimensions used in the comparison: $$
        \mathcal{K}_{ij,t} =
        \left\{
        k \;:\; x_{i,k,t} \text{ and } x_{j,k,t} \text{ are both observed}
        \right\}
    $$ $K_{ij,t} = \left| \mathcal{K}_{ij,t} \right|$ is the number of valid dimensions\

    Note: We do not impute missing data in a vector. We let the distance calculation algorithm take care of this. We have two such cases:

    -   We only have the data starting from the end of the 2017/18 season. Thus, for this season, the second and third values in the vector are not present.
    -   Also, we could have missing data for a team that was newly promoted to the league.

    In both cases, the above algorithm skips the dimension completely if the dimension is missing for either team.

4.  Next, we attribute this distance score over multiple seasons to smooth out the rivalry score, so that there aren't drastic changes in rivalry across seasons.

    $$
    R^{(s)}_{ij,t}=\alpha \, d_{ij,t}+(1 - \alpha)\, R^{(s)}_{ij,t-1}
    $$ where:

    $R^{(s)}_{ij,t}$ is the smoothed rivalry score, incorporating past rivalry\
    $\alpha \in (0,1)$ is a smoothing parameter controlling the weight placed on current versus past rivalry. After sensitivity analysis, we set $\alpha = 0.25$\

5.  Next, we calculate a continuous rivalry proximity score:

    -   First, we make the rivalry score symmetric by averaging both directions:

        $$
        R^{(\mathrm{sym})}_{ij,t}=\frac{R^{(s)}_{ij,t}+R^{(s)}_{ji,t}}{2}
        $$

    -   Next, we compute a rank-based percentile of the smoothed distance.

    $$
    P_{ij,t}=\frac{\operatorname{rank}\left(R^{(\mathrm{sym})}_{ij,t}\right)}{N_t}
    $$ where:

    $P_{ij,t}$ is the Percentile Rank for a team pair\
    $N_{t}$ is the number of team pairs in that league-season\

    -   Then, we invert it so that **when the score is closer to 1, it means that the two teams are closer rivals.**

        $$
        \text{Rivalry Proximity}_{ij,t}=1 - P_{ij,t}
        $$ {#eq-rivalry_prox}

```{r include = FALSE}
# ------4. Creating Rivalry Vector------

# ------Prepare points history-----
# Extract points table
pts_dt <- copy(dt_raw)

# normalize columns
pts_dt[, team_clean := tolower(str_squish(squad))]
pts_dt[, pts := as.numeric(pts)]

# keep only valid rows
pts_dt <- pts_dt[!is.na(pts),
                 .(country, league, season_start = season, team_clean, pts)]
pts_dt[, season_start := as.integer(season_start)]

# -------Build lagged points-------
# current season points
pts_t <- pts_dt[,.(season_start, team_clean, pts)]
setnames(pts_t, "pts", "pts_t")

# t-1
pts_t1 <- pts_dt[,.(season_start, team_clean, pts)]
pts_t1[, season_start := season_start + 1]
setnames(pts_t1, "pts", "pts_t1")

# t-2
pts_t2 <- pts_dt[,.(season_start, team_clean, pts)]
pts_t2[, season_start := season_start + 2]
setnames(pts_t2, "pts", "pts_t2")


# -----assemble base rivalry------

rival_base <- unique(
  pts_dt[, .(country, league, season_start, team_clean)]
)

# merge points
rival_base <- merge(rival_base, pts_t,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- merge(rival_base, pts_t1,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- merge(rival_base, pts_t2,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)


# ------attach elo-------
elo_dt <- matched_dt[
  , .(season_start = elo_season,
      team_clean, elo)
]

rival_base <- merge(rival_base, elo_dt,
                    by = c("season_start", "team_clean"),
                    all.x = TRUE)

rival_base <- unique(rival_base, by = c("team_clean", "season_start"))

elo_quartiles <- rival_base[
  , .(q_elo = quantile(elo, 0.25, na.rm = TRUE)),
  by = .(season_start)
]

rival_base <- merge(rival_base, elo_quartiles,
                    by = "season_start",
                    all.x = TRUE)

# rival_base[is.na(elo), elo := q_elo]


# flag seasons where lagged points are structurally missing
lag_availability <- rival_base[
  , .(
      has_pts_t1 = any(!is.na(pts_t1)),
      has_pts_t2 = any(!is.na(pts_t2))
    ),
  by = .(country, season_start)
]

rival_base <- merge(
  rival_base,
  lag_availability,
  by = c("country", "season_start"),
  all.x = TRUE
)


# -------z score------
rival_vectors <- rival_base[
  ,
  {

    # ---- identify structural seasons ----
    min_season <- min(season_start)

    # ---- impute non-structural missing ----
    pts_t1_imp <- pts_t1
    pts_t2_imp <- pts_t2

    # ---- z-score scaling (after imputation) ----
    z_t  <- as.numeric(scale(pts_t))
    z_t1 <- as.numeric(scale(pts_t1_imp))
    z_t2 <- as.numeric(scale(pts_t2_imp))
    z_e  <- as.numeric(scale(elo))

    # ---- set structural missing to 0 AFTER scaling ----
    if(season_start == 2017) {
      z_t1 <- NA_real_
      z_t2 <- NA_real_
    }
    
    if(season_start == 2018) {
      z_t2 <- NA_real_
    }

    list(
      team_clean,
      z_pts_t  = z_t,
      z_pts_t1 = z_t1,
      z_pts_t2 = z_t2,
      z_elo    = z_e
    )
  },
  by = .(country, season_start)
]

```

```{r include = FALSE}
# ------5. Calculating Rivalry Score------
rivalry_scores <- rival_vectors[
  ,
  {
    X <- as.matrix(.SD)
    teams <- team_clean
    n <- nrow(X)

    out <- vector("list", n*(n-1)/2)
    idx <- 1L

    for (i in 1:(n-1)) {
      for (j in (i+1):n) {

        xi <- X[i, ]
        xj <- X[j, ]

        # identify observed dimensions
        # ignore dimension if either team has NA (structurally missing)
        observed <- !is.na(xi) & !is.na(xj)

        k <- sum(observed)

        if (k == 0) {
          d <- NA_real_
        } else {
          d <- sqrt(sum((xi[observed] - xj[observed])^2) / k)
        }

        out[[idx]] <- list(
          team_1 = teams[i],
          team_2 = teams[j],
          rivalry_score = d
        )
        idx <- idx + 1L
      }
    }

    rbindlist(out)
  },
  by = .(country, season_start),
  .SDcols = c("z_pts_t", "z_pts_t1", "z_pts_t2", "z_elo")
]
rivalry_scores
```

```{r include = FALSE}
# ------6. Calculating Smoothed Rivalry Score------

rivalry_scores[
  ,
  `:=`(
    tA = pmin(team_1, team_2),
    tB = pmax(team_1, team_2)
  )
]

setorder(rivalry_scores, country, tA, tB, season_start)

alpha <- 0.25

rivalry_scores[
  ,
  rivalry_smooth :=
    Reduce(function(prev, curr) alpha * curr + (1 - alpha) * prev,
           rivalry_score,
           accumulate = TRUE),
  by = .(country, tA, tB)
]

rivalry_scores
```

```{r include = FALSE}
# ------6. Calculating Continuous Rivalry Proximity------

# Expand to team-centric format (same as before)
rival_long <- rbindlist(list(
  rivalry_scores[, .(country, season_start, team = tA, rival = tB, rivalry_smooth)],
  rivalry_scores[, .(country, season_start, team = tB, rival = tA, rivalry_smooth)]
))

rival_sym <- rival_long[
  rival_long,
  on = .(country, season_start,
         team = rival,
         rival = team),
  nomatch = 0,
  allow.cartesian = TRUE
][
  ,
  .(
    country,
    season_start,
    team_a = pmin(team, rival),
    team_b = pmax(team, rival),
    rivalry_symmetric = mean(c(rivalry_smooth, i.rivalry_smooth))
  ),
  by = .(country, season_start, team, rival)
][
  ,
  .SD[1],
  by = .(country, season_start, team_a, team_b)
]

rival_sym[
  ,
  rivalry_percentile := frank(rivalry_symmetric, ties.method = "average") / .N,
  by = .(country, season_start)
]

# Invert so 1 = closest rival
rival_sym[
  ,
  rivalry_proximity := 1 - rivalry_percentile,
  by = .(country, season_start)
]

rival_continuous <- rival_sym[
  ,
  .(country, season_start, team_a, team_b, rivalry_proximity)
]

rival_continuous$season_end <- rival_continuous$season_start + 1L

fwrite(rival_continuous, "data/rivalry_continuous.csv")
```

Post calculations, our rivalry look-up table looks like @tbl-riv

```{r}
#| tbl-cap: "Top 5 Rivalry Proximity Scores"
#| label: tbl-riv

kable(head(rival_continuous[order(rivalry_proximity, decreasing = TRUE)], n = 5), booktabs = TRUE) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),
    font_size = 9
  )
```

## Mapping

We have gathered league table, transfers and ELO from different sources. We have merged league table and ELO to create the rivalry vector. But we still need to make sure these team names match with the ones in transfers. Thus, before we can proceed with the analysis, we use fuzzy matching to match team names across tables. The mismatched teams after fuzzy match are stored in a csv file in which we then manually add the correct team name. This file is then used to complete the mapping.

```{r include = FALSE}
# ------7. Mapping Team names------

## ===================== PATHS ===================== ##
rivals_path    <- "data/rivalry_continuous.csv"
transfers_path <- "data/top5_league_transfers.csv"
league_path <- "data/league_table.csv"
out_dir     <- "data"

## ===================== 1. LOAD DATA ===================== ##
league <- fread(league_path)
xf     <- fread(transfers_path)

## ===================== 2. BUILD CANONICAL TEAM LIST ===================== ##
# League table defines truth
league[, team_clean := tolower(str_squish(squad))]
league[, country := as.character(country)]

canon_teams <- unique(
  league[, .(country, team_clean)]
)

## ===================== 3. EXTRACT RAW TRANSFER TEAM NAMES ===================== ##
xf[, buyer_raw  := tolower(str_squish(team_name))]
xf[, seller_raw := tolower(str_squish(club_2))]
xf[, country    := as.character(country)]

raw_teams <- unique(rbindlist(list(
  xf[, .(country = country, raw = buyer_raw)],
  xf[, .(country = country_2, raw = seller_raw)]
)))

valid_canon <- canon_teams[, .(country, team_clean)]

raw_teams <- raw_teams[
  country %in% valid_canon$country
]

normalize <- function(x) {
  x |>
    tolower() |>
    str_replace_all("&", "and") |>
    str_replace_all("[^a-z ]", "") |>
    str_squish()
}

valid_canon[, team_norm := normalize(team_clean)]
raw_teams[, raw_norm := normalize(raw)]


## ===================== 4. FUZZY MATCH (RAW → CANONICAL) ===================== ##
team_map <- stringdist_inner_join(
  raw_teams,
  valid_canon,
  by = c("raw" = "team_clean"),
  method = "jw",
  max_dist = 0.2,
  distance_col = "dist"
)

# IMPORTANT: force data.table
setDT(team_map)

## ===================== 5. FIX country.x / country.y ===================== ##
# Keep only same-country matches
team_map <- team_map[country.x == country.y]

# Clean column names
setnames(team_map, "country.x", "country")
team_map[, country.y := NULL]
team_map[, country.dist := NULL]
team_map[, raw.dist := NULL]

# Keep best match per (country, raw)
setorder(team_map, country, raw, dist)
team_map <- team_map[, .SD[1], by = .(country, raw)]
```

```{r include = FALSE}
riv <- fread(rivals_path)   # rivalry_continuous.csv
manual_map <- fread("data/unmatched_transfer_teams.csv")

# Clean rivalry team names
riv[, team_a_clean := tolower(str_squish(team_a))]
riv[, team_b_clean := tolower(str_squish(team_b))]
riv[, season := as.integer(season_end)]

# Keep only needed columns
riv_unique <- unique(
  riv[, .(country, season, team_a_clean, team_b_clean, rivalry_proximity)]
)

# ===================== 2. CLEAN TRANSFER TEAM NAMES ===================== #

team_map_final <- unique(rbind(
  team_map[, .(country, raw, team_clean)],
  manual_map
))

team_map_final[, team_clean := tolower(str_squish(team_clean))]

# Buyer clean
xf <- merge(
  xf,
  team_map_final,
  by.x = c("country", "buyer_raw"),
  by.y = c("country", "raw"),
  all.x = TRUE
)
setnames(xf, "team_clean", "buyer_clean")

xf <- merge(
  xf,
  team_map_final,
  by.x = c("country_2", "seller_raw"),
  by.y = c("country", "raw"),
  all.x = TRUE
)
setnames(xf, "team_clean", "seller_clean")

# there are still mismatches for country_2 (our two tables identify teams like Monaco in
# different countries. In one table, country_2 for monaco is Monaco. 
# In the other, it is France. After veriyfing the data, we can safely 
# use the same name from seller_raw toseller_clean for such cases)

xf[is.na(seller_clean), seller_clean := seller_raw]

# Remove duplicate columns from merges
xf <- xf[, !duplicated(names(xf), fromLast = TRUE), with = FALSE]

# ===================== 3. MAKE TRANSFER PAIRS UNDIRECTED ===================== #

# Create ordered pair to match rivalry file
xf[, team_low  := pmin(buyer_clean, seller_clean)]
xf[, team_high := pmax(buyer_clean, seller_clean)]

# ===================== 4. MERGE CONTINUOUS RIVALRY ===================== #

xf <- merge(
  xf,
  riv_unique,
  by.x = c("country", "season", "team_low", "team_high"),
  by.y = c("country", "season", "team_a_clean", "team_b_clean"),
  all.x = TRUE
)

# ===================== 5. HANDLE CROSS-LEAGUE TRANSFERS ===================== #

# If seller is from a different league (country_2 != country),
# rivalry should be NA explicitly

xf[country_2 != country, rivalry_proximity := NA_real_]

# ===================== 6. CLEANUP ===================== #

xf[, c("team_low", "team_high") := NULL]

xf[, pos_clean := tolower(str_squish(player_position))]
```

Once the mapping is done, we check again to ensure the data are clean and there are no mismatched sellers. We limit our check for mismatches to sellers from just the top 5 leagues (we don't have any data for sellers outside the top 5 leagues so these would get falsely flagged as mismatches).

```{r include = FALSE}
# ------8. Check Seller name mismatch------
setDT(league_table)
setDT(xf)

## ---------- NORMALIZATION FUNCTION ----------
normalize <- function(x) {
  x |>
    tolower() |>
    str_replace_all("&", "and") |>
    str_replace_all("[^a-z0-9 ]", "") |>
    str_squish()
}

## ---------- NORMALIZE TEAM NAMES ----------
league_table[, squad_norm := normalize(squad)]
xf[, seller_norm := normalize(seller_clean)]

## ---------- REMOVE DUPLICATES ----------
league_teams <- unique(
  league_table[, .(league_name, squad, squad_norm)]
)

seller_teams <- unique(
  xf[, .(league_name = league_2, seller_clean, seller_norm)]
)

## ---------- FIND MISMATCHES ----------
mismatches <- seller_teams[league_name %in% top5_leagues][
  !league_teams,
  on = .(league_name, seller_norm = squad_norm)
]
```

```{r}
#| tbl-cap: "Seller team name mismatches"
#| label: tbl-seller-miss

kable(head(mismatches, n = 5), booktabs = TRUE) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),
    font_size = 9
  )
```

From @tbl-seller-miss we can clearly see that these are mostly teams mismatched to the league (Ex: Salzburg is a team from the Austrian Bundesliga, but is matched to the German Bundesliga). These erroneous entries can be removed.

```{r include = FALSE}
xf <- xf[
  !mismatches,
  on = .(league_2 = league_name, seller_norm)
]
```

Similarly, when we checked for buyers, no mismatches were found.

```{r include=FALSE}
# ------9. Check buyer name mismatch------
xf[, buyer_norm := normalize(buyer_clean)]
buyer_teams <- unique(
  xf[, .(league_name = league, buyer_clean, buyer_norm)]
)

## ---------- FIND MISMATCHES ----------
mismatches <- buyer_teams[
  !league_teams,
  on = .(league_name, buyer_norm = squad_norm)
]
```

```{r include = FALSE}
fwrite(xf, "data/top5_league_transfers_clean.csv")
```

# Choice Set Construction

Now, our data are cleaned and ready. Next, we construct choice datasets tailored to the analyses. Each choice set contains:

-   Focal transfer: The observed transfer. The buying team is known as the focal team, and its league is the focal league.

-   Alternative pool: All the other options that the buying team has for the focal transfer. For each focal transfer, the feasible alternatives pool is constructed based on these conditions:

    -   The transfer must be in the same window as the focal transfer (this is done using the focal season).

    -   The alternative player must play in the same position as the focal player.

    -   The transfer fee of the alternative must be comparable to the transfer fee of the focal player. (If the focal player was a free transfer, then a player with an 80 million transfer fee cannot be considered an alternative). Therefore, we restrict the choice set to players whose transfer fees lie within a symmetric bandwidth around the focal player’s fee in logarithmic terms.

        $$
        \left| \log(\text{fee}_j) - \log(\text{fee}_i) \right| \leq h
        $$

        where:

        $fee_i$: transfer fee of the focal player\
        $fee_j$: transfer fee of the alternative player\
        $h$: the bandwidth parameter governing the admissible range of transfer fees (we set h = 1 after sensitivity analysis)\

These choice sets are identified by the `choice_id`. For each choice set, the transfer with chosen = 1 is the focal transfer. The other transfers in the choice set are marked with chosen = 0 and are the alternatives for the focal transfer. We do not consider transfers with no alternative options.

These are the independent variables that we use in our analysis:

-   `player_age` :age of the player during the transfer.

-   `minutes_played` :the number of confirmed minutes the player has played in the previous season.

-   `nat_match`: denotes if a player (focal/ alternative) belongs to the focal country.

-   `goals`: the number of goals a player scored the previous season.

-   `log_fee`: The transfer fee of the player in log scale

-   `rivalry_proximity`: The rivalry score we calculated based on @eq-rivalry_prox. We compute this score for the focal team against the alternative transfer's selling team (Sometimes referred to as just rivalry in this article).

The dependent variable is `chosen`. It is a binary variable, where the value is 1 if the player was chosen, 0 if not.

## Choice Dataset 1

Here, **in the focal group we include only permanent arrivals to the top 5 leagues** (sellers can be outside the top 5 leagues). **The buyers of the alternatives are from the same league as the focal buyer** (alternative league = focal league). We consider loan options in the alternative pool; not in the focal group.

```{r include = FALSE}
transfers <- read_csv("data/top5_league_transfers_clean.csv")
rivalry_lookup <- read_csv("data/rivalry_continuous.csv")

transfers <- transfers %>%
  mutate(
    transfer_id = row_number(),
    log_fee = log(transfer_fee + 1)
  )

band_width <- 1
set.seed(123)
K <- 10

choice_data_1 <- transfers %>%
  filter(
    is_loan == FALSE
  ) %>%
  rowwise() %>%
  do({
    
    current <- .
    
    focal_log_fee  <- current$log_fee[[1]]
    focal_season   <- current$season[[1]]
    focal_pos      <- current$pos_clean[[1]]
    focal_country  <- current$country[[1]]
    focal_league  <- current$league[[1]]
    focal_id       <- current$transfer_id[[1]]
    
    pool_all <- transfers %>%
      filter(
        season == focal_season,
        pos_clean == focal_pos,
        league == focal_league,
        transfer_id != focal_id,
        !is.na(log_fee),
        abs(log_fee - focal_log_fee) <= band_width
      )
    
    # Randomly sample up to K alternatives
    if (nrow(pool_all) > K) {
      pool_sample <- pool_all %>% slice_sample(n = K)
    } else {
      pool_sample <- pool_all
    }
    
    pool <- bind_rows(pool_sample, current)
    
    #recompute rivalry score
    focal_buyer_clean <- current$buyer_clean[[1]]
    pool <- pool %>%
    mutate(
        riv_low  = pmin(focal_buyer_clean, seller_clean),
        riv_high = pmax(focal_buyer_clean, seller_clean)
      ) %>%
      left_join(
        rivalry_lookup %>%
          mutate(team_a = tolower(str_squish(team_a)),
                 team_b = tolower(str_squish(team_b))) %>%
          filter(country == focal_country),
        by = c("riv_low" = "team_a",
               "riv_high" = "team_b",
               "season" = "season_end")
      ) %>%
      mutate(
        rivalry_proximity = coalesce(rivalry_proximity.y, 0)
      ) %>%
      select(-rivalry_proximity.x, -riv_low, -riv_high)
    
    # compute nationality match and mark chosen transfers
    pool %>%
      mutate(
        choice_id = focal_id,
        nat_match = ifelse(player_nationality == focal_country, 1, 0),
        chosen = ifelse(transfer_id == focal_id, 1, 0)
      )
    
  }) %>%
  ungroup()

choice_data_1 <- choice_data_1 %>%
  mutate(
    rivalry_proximity = ifelse(is.na(rivalry_proximity), 0, rivalry_proximity)
  )

choice_data_1$rivalry_proximity.y <- NULL
choice_data_1$country.y <- NULL
choice_data_1 <- rename(choice_data_1, country = country.x)
```

```{r include = FALSE}
#remove choice sets where there are no alternatives
setDT(choice_data_1)
choice_data_1 <- choice_data_1[
  choice_id %in% choice_data_1[, .N, by = choice_id][N > 1, choice_id]
]

#we save the data table with all columns for manual verification purposes
fwrite(choice_data_1,"data/choice_data_1.csv")

choice_dataset_1 <- choice_data_1 %>%
  select(
    chosen,
    player_age,
    minutes_played,
    nat_match,
    goals,
    log_fee,
    rivalry_proximity,
    choice_id
  )
```

@tbl-cds1 shows a sample of this choice dataset.

```{r}
#| tbl-cap: "Choice Dataset 1"
#| label: tbl-cds1

kable(filter(choice_dataset_1, choice_id == 138), booktabs = TRUE) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),
    font_size = 9
  )
```

## Choice Dataset 2

This is the same as the previous dataset, but here we **also include the transfers where the seller is from the focal league**. Thus, **the player should either leave the focal league or join it**.

Thus, for the same choice set as in @tbl-cds1, our new method of construction gives extra alternatives, as shown in @tbl-cds2.

```{r include = FALSE}
set.seed(123)
band_width <- 1
K <- 10

choice_data_2 <- transfers %>%
  filter(
    is_loan == FALSE
  ) %>%
  rowwise() %>%
  do({
    
    current <- .
    
    focal_log_fee  <- current$log_fee[[1]]
    focal_season   <- current$season[[1]]
    focal_country  <- current$country[[1]]
    focal_league   <- current$league[[1]]
    focal_pos      <- current$pos_clean[[1]]
    focal_id       <- current$transfer_id[[1]]
    
    # Candidate alternatives
    pool_all <- transfers %>%
      filter(
        season == focal_season,
        league == focal_league | league_2 == focal_league,
        pos_clean == focal_pos,
        transfer_id != focal_id,
        !is.na(log_fee),
        abs(log_fee - focal_log_fee) <= band_width
      )
    
    # Randomly sample up to K alternatives
    if (nrow(pool_all) > K) {
      pool_sample <- pool_all %>% slice_sample(n = K)
    } else {
      pool_sample <- pool_all
    }
    
    # Add chosen transfer back
    pool <- bind_rows(pool_sample, current)
    
    #recompute rivalry score
    focal_buyer_clean <- current$buyer_clean[[1]]
    pool <- pool %>%
    mutate(
        riv_low  = pmin(focal_buyer_clean, seller_clean),
        riv_high = pmax(focal_buyer_clean, seller_clean)
      ) %>%
      left_join(
        rivalry_lookup %>%
          mutate(team_a = tolower(str_squish(team_a)),
                 team_b = tolower(str_squish(team_b))) %>%
          filter(country == focal_country),
        by = c("riv_low" = "team_a",
               "riv_high" = "team_b",
               "season" = "season_end")
      ) %>%
      mutate(
        rivalry_proximity = coalesce(rivalry_proximity.y, 0)
      ) %>%
      select(-rivalry_proximity.x, -riv_low, -riv_high)
    
    pool %>%
      mutate(
        choice_id = focal_id,
        nat_match = ifelse(player_nationality == focal_country, 1, 0),
        chosen = ifelse(transfer_id == focal_id, 1, 0)
      )
    
  }) %>%
  ungroup()

choice_data_2 <- choice_data_2 %>%
  mutate(
    rivalry_proximity = ifelse(is.na(rivalry_proximity), 0, rivalry_proximity)
  )

setDT(choice_data_2)
choice_data_2 <- choice_data_2[
  choice_id %in% choice_data_2[, .N, by = choice_id][N > 1, choice_id]
]

fwrite(choice_data_2, "data/choice_data_2.csv")

choice_dataset_2 <- choice_data_2 %>%
  select(
    chosen,
    player_age,
    minutes_played,
    nat_match,
    goals,
    log_fee,
    rivalry_proximity,
    choice_id
  )
```

```{r}
#| tbl-cap: "Choice Dataset 2"
#| label: tbl-cds2

kable(filter(choice_dataset_2, choice_id == 138), booktabs = TRUE) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),
    font_size = 9
  )
```

## Choice Dataset 3

In choice dataset 2 we include alternatives where the seller is from the focal league (but the buyer may be from a different league). For these alternatives, since the seller is from the focal league, the rivalry between the alternative seller and the focal buyer is non-zero. But there are also cases where a team buys players from outside the focal league. Here, the rivalry is always zero between the focal buyer and sellers outside the focal league.

Thus, any effect that rivalry_proximity has may therefore not reflect genuine preference/avoidance. Instead, it could be picking up: "chosen transfers tend to have cross-league sellers (rivalry = 0) while unchosen within-league-seller alternatives have rivalry \> 0." The model would be learning "same-league seller pair vs. cross-league seller pair" rather than "rivalry matters."

To test this, we now construct a choice dataset by **excluding all transfers from outside the focal league** (thus **all focal and alternative transfers are within the same league**). If the result is the same as the model designed using the previous choice set, then the effect can be considered genuine.

```{r include = FALSE}
set.seed(123)
band_width <- 1
K <- 10

choice_data_3 <- transfers %>%
  filter(
    is_loan == FALSE,
    league_2 == league
  ) %>%
  rowwise() %>%
  do({
    
    current <- .
    
    focal_log_fee  <- current$log_fee[[1]]
    focal_season   <- current$season[[1]]
    focal_country  <- current$country[[1]]
    focal_league   <- current$league[[1]]
    focal_pos      <- current$pos_clean[[1]]
    focal_id       <- current$transfer_id[[1]]
    
    # Candidate alternatives
    pool_all <- transfers %>%
      filter(
        season == focal_season,
        league == focal_league & league_2 == focal_league,
        pos_clean == focal_pos,
        transfer_id != focal_id,
        !is.na(log_fee),
        abs(log_fee - focal_log_fee) <= band_width
      )
    
    # Randomly sample up to K alternatives
    if (nrow(pool_all) > K) {
      pool_sample <- pool_all %>% slice_sample(n = K)
    } else {
      pool_sample <- pool_all
    }
    
    # Add chosen transfer back
    pool <- bind_rows(pool_sample, current)
    
    #recompute rivalry score
    focal_buyer_clean <- current$buyer_clean[[1]]
    pool <- pool %>%
    mutate(
        riv_low  = pmin(focal_buyer_clean, seller_clean),
        riv_high = pmax(focal_buyer_clean, seller_clean)
      ) %>%
      left_join(
        rivalry_lookup %>%
          mutate(team_a = tolower(str_squish(team_a)),
                 team_b = tolower(str_squish(team_b))) %>%
          filter(country == focal_country),
        by = c("riv_low" = "team_a",
               "riv_high" = "team_b",
               "season" = "season_end")
      ) %>%
      mutate(
        rivalry_proximity = coalesce(rivalry_proximity.y, 0)
      ) %>%
      select(-rivalry_proximity.x, -riv_low, -riv_high)
    
    pool %>%
      mutate(
        choice_id = focal_id,
        nat_match = ifelse(player_nationality == focal_country, 1, 0),
        chosen = ifelse(transfer_id == focal_id, 1, 0)
      )
    
  }) %>%
  ungroup()

choice_data_3 <- choice_data_3 %>%
  mutate(
    rivalry_proximity = ifelse(is.na(rivalry_proximity), 0, rivalry_proximity)
  )

setDT(choice_data_3)
choice_data_3 <- choice_data_3[
  choice_id %in% choice_data_3[, .N, by = choice_id][N > 1, choice_id]
]

fwrite(choice_data_3, "data/choice_data_3.csv")

choice_dataset_3 <- choice_data_3 %>%
  select(
    chosen,
    player_age,
    minutes_played,
    nat_match,
    goals,
    log_fee,
    rivalry_proximity,
    choice_id
  )
```

```{r}
#| tbl-cap: "Choice Dataset 3"
#| label: tbl-cds3

#there are NA values that we don't handle; the model will simply ignore such values
kable(filter(choice_dataset_3, choice_id == 138), booktabs = TRUE) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"),
    font_size = 9
  )
```

# Model And Results

For our analysis, we need to model why a chosen player was preferred over all other feasible alternatives. Thus, we use the **Conditional Multinomial Logit Model.**

$$
  P_{ij} =\frac{\exp(X_{ij}\beta)}{\sum_{k \in C_i} \exp(X_{ik}\beta)}
  $$

where:

$i$: choice situation (choice_id)\
$C_i$: Choice set for $i$\
$j$: jth alternative in $C_i$\
$P_{ij}$: probability that alternative $j$ is chosen from choice set $C_i$\
$X_{ij}$: observed characteristics (independent variables)\

For the choice datasets that we have constructed so far we estimate a model for each set:

## Model 1

Model 1 uses choice dataset 1 (permanent arrivals to the top 5 leagues in the focal group and the buyers of the alternatives are from the same league as the focal buyer)

```{r echo=TRUE}
model_1 <- clogit(
  chosen ~ player_age +
    minutes_played + 
    nat_match +
    goals +
    log_fee + 
    rivalry_proximity + 
    strata(choice_id),
  data = choice_dataset_1
)
```

```{r}
#| tbl-cap: "Model 1 Summary"
#| label: tbl-m1

custom <- tidy(model_1) %>%
  mutate(
    odds_ratio = exp(estimate),
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value < 0.1 ~ "+",
      TRUE ~ ""
    ),
    coef_star = paste0(round(estimate, 3), stars),
    or_round  = round(odds_ratio, 3),
    std.error = round(std.error, 3)
  ) %>%
  select(
    Variable = term,
    Coef = coef_star,
    `Exp(Coef)` = or_round,
    SE = std.error
  ) %>%
  mutate(across(everything(), as.character))

# GOF stats (match column names!)
gof <- data.frame(
  Variable = c("Concordance", "Log-Likelihood", "p-value"),
  Coef = c(round(summary(model_1)$concordance["C"] * 100, 2),
           round(summary(model_1)$logtest["test"], 2),
           ifelse(summary(model_1)$logtest["pvalue"] < 0.05, "<0.05", ">0.05"))
) %>%
  mutate(
    Coef = as.character(Coef),
    `Exp(Coef)` = "",
    SE = ""
  )

# Combine
final_table <- bind_rows(custom, gof)

# Display
kable(final_table)
```

Based on the log likelihood test score and low p-value, this model is statistically significant and provides a reasonable fit. The concordance, although low, is still \> 50%, suggesting our model performs better than random selection. So we can look into the coefficients to find out what the model has learned.

Using @tbl-m1 we can investigate the independent variables. For rivalry proximity, we have:

```{r include = FALSE}
# setting dynamic variables
format_p <- function(p) {
  if (p < 0.001) {
    "< 0.001"
  } else {
    paste0("≈ ", round(p, 3))
  }
}

#for rivalry
riv_coef <- coef(model_1)["rivalry_proximity"]
riv_exp <- exp(coef(model_1)["rivalry_proximity"])
riv_perc <- (riv_exp - 1) * 100
riv_dir <- ifelse(riv_perc > 0, "increase", "decrease")
riv_pval <- summary(model_1)$coefficients["rivalry_proximity", "Pr(>|z|)"]
riv_stat <- ifelse(riv_pval < 0.05, "statistically significant (low p-value)", "insignificant (p-value > 0.05)")
sig_riv <- riv_pval < 0.05
```

::: {.callout-note icon="false" appearance="simple"}
coef **≈** `r round(riv_coef, 3)`.\
exp(coef) **≈** `r round(riv_exp, 3)`.\
p `r format_p(riv_pval)`
:::

This effect is `r riv_stat`.`r if (sig_riv) paste0("A one-unit increase in rivalry proximity leads to a **",  round(abs(riv_perc), 1), "% ", riv_dir, " in the odds of selection**.")`

Now we look at other factors that might affect the choice of a team when looking for a player. When we check for player_age:

```{r include = FALSE}
# setting dynamic variables

#for player_age
age_coef <- coef(model_1)["player_age"]
age_exp <- exp(coef(model_1)["player_age"])
age_perc <- (age_exp - 1) * 100
age_dir <- ifelse(age_perc > 0, "increase", "decrease")
age_pval <- summary(model_1)$coefficients["player_age", "Pr(>|z|)"]
age_stat <- ifelse(age_pval < 0.05, "statistically significant (low p-value)", "insignificant (p-value > 0.05)")
sig_age <- age_pval < 0.05
```

::: {.callout-note icon="false" appearance="simple"}
coef **≈** `r round(age_coef, 3)`.\
exp(coef) **≈** `r round(age_exp, 3)`.\
p `r format_p(age_pval)`
:::

This is `r age_stat`. `r if (sig_age) paste0(   "A one-year increase in a player's age leads to a **",   round(abs(age_perc), 1), "% ", age_dir,   " in the odds of selection.**",   if (age_perc > 0)     " Although this isn't the absolute probability of being signed, it means that clubs value players with more experience (proxied by age) conditioned on fee." )`

We have another significant variable in minutes_played:

```{r include = FALSE}
# setting dynamic variables

#for minutes_played
mins_coef <- coef(model_1)["minutes_played"]
mins_exp <- exp(coef(model_1)["minutes_played"])
mins_perc <- (mins_exp - 1) * 100
mins_dir <- ifelse(mins_perc > 0, "increase", "decrease")
mins_pval <- summary(model_1)$coefficients["minutes_played", "Pr(>|z|)"]
mins_stat <- ifelse(mins_pval < 0.05, "statistically significant (low p-value)", "insignificant (p-value > 0.05)")
sig_mins <- mins_pval < 0.05
rnd_mins_perc <- round(abs(mins_perc), 1)
```

::: {.callout-note icon="false" appearance="simple"}
coef **≈** `r round(mins_coef, 3)`.\
exp(coef) **≈** `r round(mins_exp, 3)`.\
p `r format_p(mins_pval)`
:::

This is `r mins_stat`.`r if (sig_mins & rnd_mins_perc != 0) paste0(" A one minute increase in a player's playing time leads to a **",  rnd_mins_perc, "% ", mins_dir, " in the odds of the player being chosen over the alternatives.**")` `r if (sig_mins & rnd_mins_perc == 0) paste0(" But **the effect is negligible.**")`

There are other factors such as goals, nationality match and transfer fee, but when we restrict the focal and alternative pool to only include players that arrive to the focal league, these factors are insignificant. This might change when we also include players who left the focal league in the same window.

Note: all effects on all models include the condition that the fee range must be similar for the focal and its alternatives. Thus, the effect is not absolute. It is conditional on price.

## Model 2

Model 2 uses choice dataset 2 (the player should either leave the focal league or join it). This gives us the following result:

```{r echo = TRUE}
model_2 <- clogit(
  chosen ~ player_age +
    minutes_played + nat_match +
    goals +
    log_fee + 
    rivalry_proximity + 
    strata(choice_id),
  data = choice_dataset_2
)
```

```{r}
#| tbl-cap: "Model 2 Summary"
#| label: tbl-m2

custom_m2 <- tidy(model_2) %>%
  mutate(
    odds_ratio = exp(estimate),
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value < 0.1 ~ "+",
      TRUE ~ ""
    ),
    coef_star = paste0(round(estimate, 3), stars),
    or_round  = round(odds_ratio, 3),
    std.error = round(std.error, 3)
  ) %>%
  select(
    Variable = term,
    Coef = coef_star,
    `Exp(Coef)` = or_round,
    SE = std.error
  ) %>%
  mutate(across(everything(), as.character))

# GOF stats (match column names!)
gof_m2 <- data.frame(
  Variable = c("Concordance", "Log-Likelihood", "p-value"),
  Coef = c(round(summary(model_2)$concordance["C"] * 100, 2),
           round(summary(model_2)$logtest["test"], 2),
           ifelse(summary(model_2)$logtest["pvalue"] < 0.05, "<0.05", ">0.05"))
) %>%
  mutate(
    Coef = as.character(Coef),
    `Exp(Coef)` = "",
    SE = ""
  )

# Combine
final_table_m2 <- bind_rows(custom_m2, gof_m2)

# Display
kable(final_table_m2)
```

@tbl-m2 shows that model 2 has concordance and likelihood ratio test score both higher than model 1. This suggests that the **second model is a better fit and has more discriminative power compared to model 1.** Now let us look into the variables of interest. Our main focus is on rivalry_proximity:

```{r include = FALSE}

#for rivalry
riv_coef_m2 <- coef(model_2)["rivalry_proximity"]
riv_exp_m2 <- exp(coef(model_2)["rivalry_proximity"])
riv_perc_m2 <- (riv_exp_m2 - 1) * 100
riv_dir_m2 <- ifelse(riv_perc_m2 > 0, "increase", "decrease")
riv_pval_m2 <- summary(model_2)$coefficients["rivalry_proximity", "Pr(>|z|)"]
riv_stat_m2 <- ifelse(riv_pval_m2 < 0.05, "statistically significant (low p-value)", "insignificant (p-value > 0.05)")
sig_riv_m2 <- riv_pval_m2 < 0.05
```

::: {.callout-note icon="false" appearance="simple"}
coef **≈** `r round(riv_coef_m2, 3)`.\
exp(coef) **≈** `r round(riv_exp_m2, 3)`.\
p `r format_p(riv_pval_m2)`
:::

This effect is `r riv_stat_m2`.`r if (sig_riv_m2) paste0(" A one-unit increase in rivalry proximity leads to a **",  round(abs(riv_perc_m2), 1), "% ", riv_dir_m2, " in the odds of selection**.")`

`r if (sig_riv_m2) paste0("We look more into this in the next section, where we validate this outcome.")`

Next is player_age:

```{r include = FALSE}
# setting dynamic variables

#for player_age
age_coef_m2 <- coef(model_2)["player_age"]
age_exp_m2 <- exp(coef(model_2)["player_age"])
age_perc_m2 <- (age_exp_m2 - 1) * 100
age_dir_m2 <- ifelse(age_perc_m2 > 0, "increase", "decrease")
age_pval_m2 <- summary(model_2)$coefficients["player_age", "Pr(>|z|)"]
age_stat_m2 <- ifelse(age_pval_m2 < 0.05, "statistically significant (low p-value)", "insignificant (p-value > 0.05)")
sig_age_m2 <- age_pval_m2 < 0.05
```

::: {.callout-note icon="false" appearance="simple"}
coef **≈** `r round(age_coef_m2, 3)`.\
exp(coef) **≈** `r round(age_exp_m2, 3)`.\
p `r format_p(age_pval_m2)`
:::

This is `r age_stat_m2`. `r if (sig_age_m2) paste0(   "A one-year increase in a player's age leads to a **",   round(abs(age_perc_m2), 1), "% ", age_dir_m2,   " in the odds of selection.**")`

Now, for nationality match (nat_match):

```{r include = FALSE}
# setting dynamic variables

#for nat_match
nat_coef_m2 <- coef(model_2)["nat_match"]
nat_exp_m2 <- exp(coef(model_2)["nat_match"])
nat_perc_m2 <- (nat_exp_m2 - 1) * 100
nat_dir_m2 <- ifelse(nat_perc_m2 > 0, "increase", "decrease")
nat_pval_m2 <- summary(model_2)$coefficients["nat_match", "Pr(>|z|)"]
nat_stat_m2 <- ifelse(nat_pval_m2 < 0.05, "statistically significant (low p-value)", "insignificant (p-value > 0.05)")
sig_nat_m2 <- nat_pval_m2 < 0.05
```

::: {.callout-note icon="false" appearance="simple"}
coef **≈** `r round(nat_coef_m2, 3)`.\
exp(coef) **≈** `r round(nat_exp_m2, 3)`.\
p `r format_p(nat_pval_m2)`
:::

This is `r nat_stat_m2`. `r if (sig_nat_m2) paste0(   "A one-unit increase in nationality match leads to a **",   round(abs(nat_perc_m2), 1), "% ", nat_dir_m2,   " in the odds of selection.**")`

```{r include = FALSE}
# setting dynamic variables

#for minutes_played
mins_coef_m2 <- coef(model_2)["minutes_played"]
mins_exp_m2 <- exp(coef(model_2)["minutes_played"])
mins_perc_m2 <- (mins_exp_m2 - 1) * 100
mins_dir_m2 <- ifelse(mins_perc_m2 > 0, "increase", "decrease")
mins_pval_m2 <- summary(model_2)$coefficients["minutes_played", "Pr(>|z|)"]
mins_stat_m2 <- ifelse(mins_pval_m2 < 0.05, "statistically significant (low p-value)", "insignificant (p-value > 0.05)")
sig_mins_m2 <- mins_pval_m2 < 0.05
rnd_mins_perc_m2 <- round(abs(mins_perc_m2), 1)
```

`r if (sig_mins_m2 & rnd_mins_perc_m2 != 0) paste0(   "A one minute increase in the time a player played last season leads to a **",   rnd_mins_perc_m2, "% ", mins_dir_m2,   " in the odds of selection.**")`

`r if ((sig_mins & rnd_mins_perc == 0) & (sig_mins_m2) & rnd_mins_perc_m2 == 0) paste0("Just like model 1, model 2 gives a **significant yet negligible output for minutes_played.**")`

## Model 3

For the third model, we use choice dataset 3, which is used to verify the robustness of the outputs of model 2. From model 2, we found out that rivalry has a negative effect on a transfer. As explained earlier in [Choice Dataset 3], this could be a result of cross-league transfers having zero rivalry. If model 3 gives a significant negative coefficient for rivalry_proximity, then the effect is genuine. We compare the two models:

```{r echo = TRUE}
model_3 <- clogit(
  chosen ~ player_age +
    minutes_played + nat_match +
    goals +
    log_fee + 
    rivalry_proximity + 
    strata(choice_id),
  data = choice_dataset_3
)
```

```{r}
#dynamic variables 
#for model 3 rivalry
riv_exp_m3 <- exp(coef(model_3)["rivalry_proximity"])
riv_perc_m3 <- (riv_exp_m3 - 1) * 100
```

```{r}
#| tbl-cap: "Model 2 vs Model 3"
#| label: tbl-m2m3
#| tbl-pos: H

# function to add stars
add_stars <- function(p) {
  case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    p < 0.1   ~ "+",
    TRUE ~ ""
  )
}

# tidy models
df <- bind_rows(
  tidy(model_2) %>% mutate(model = "Model 2 Coef"),
  tidy(model_3) %>% mutate(model = "Model 3 Coef")
) %>%
  mutate(
    stars = add_stars(p.value),
    coef_star = sprintf("%.3f%s", estimate, stars)
)
df_wide <- df %>%
  select(term, model, coef_star) %>%
  pivot_wider(names_from = model, values_from = coef_star)

kable(df_wide, booktabs = TRUE, align = "lcc") %>%
  kable_styling(latex_options = c("hold_position"))
```

We can see in @tbl-m2m3 that although the effect of rivalry is significant here as well, it has the opposite effect on a transfer choice to what model 2 depicted. Here, we get a positive effect of `r paste0(round(abs(riv_perc_m3), 1))`%. This means that **the effect we saw from rivalry_proximity on a player being chosen in model 2 is indeed driven by the fact that cross-seller transfers have zero rivalry effect**. Therefore, this effect can be disregarded.

# Discussion & Conclusion

Overall, we find that a **player's age is an important driving factor**, and that **teams value a player slightly more if they have more experience for the same price range**. Also, clubs consider player availability (minutes played), but it doesn't seem to have a great deal of effect on the choice. Additionally, when the teams consider players who either leave or join their league as options, they seem to **value a player who is familiar with the country more**.

Other factors such as goals and rivalry do not appear to matter significantly, as the negative effect from rivalry when the choice set is included both arrivals and departures from the league seems to be merely an artifact arising from the fact that teams from different leagues don't consider themselves as rivals. Assuming that teams consider only alternatives that fit their budget, **cost doesn't seem to be the decisive factor as long as the players fit the other criteria**.
